---
title: "Bay Scallop Larve - D Stage Abnormality Rates"
author: "SamGurr"
date: "12/1/2024"
output: html_document
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_F3_Larvae_OA/RAnalysis") # Sam's
knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Airradians_F3_Larvae_OA/RAnalysis") # Sam's

```

```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(lmtest)
```

### Load Data as df
```{r}
#Here we are assing our data fram the name "df" so that we do not have to type out the full file name everytime we want to call it.
df<-read.csv("Data/Physiology/Dstage.data.csv", header = T) %>% 
  # change to numeric for the DV and as factor for IDVs
  dplyr::mutate(Percent.deformed = as.numeric(Percent.deformed),
                Rep              = as.factor(Rep),
                Year             = as.factor(Year),
                Run              = as.factor(Run)) %>% 
  
  # rename the variables for consistancy using 'case_when'
  dplyr::mutate(Parent = case_when(Parent ==  "LOW" ~ "Low pCO2", 
                                   Parent ==  "MODERATE" ~ "Moderate pCO2", 
                                   Parent ==  "HIGH" ~ "High pCO2")) %>% 
  dplyr::mutate(Treatment = case_when(Treatment ==  "Low OA" ~ "Low pCO2", 
                                      Treatment ==  "Moderate OA" ~ "Moderate pCO2", 
                                      Treatment ==  "High OA" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parent = as.factor(forcats::fct_relevel(Parent, 
                                                        c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(Treatment = as.factor(forcats::fct_relevel(Treatment, 
                                                           c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parent,
                Larvae_pCO2 = Treatment,
                Expeirment  = Run) # done 


#to look at columns names of your data frame
head(df)

#Sturcture of the data - to check that variables are properly assigned to facotr or variable
str(df)
```

# About data 

### *NOTE* target data for this repositor of generation 3 in Experiment #6!!! (F3)

### Summary: 

  - (1) F1 larvae under three OA levels, *no grow out*

  - (2) F1 larvae under three OA levels, *grown out to adult stage*

  - (3) F2 larvae full cross 3 x 3 parent x offspring OA, *no grow out*

  - (4) F2 larvae under two OA levels, *grown out to adult stage*

  - (5) F2 larvae under 3 x 3 parent x offspring OA, *no grow out*

  - (6) **F3 larvae under 3 x 3 parent x offsring OA**, *no grow out past juvenile takedown of the experiment*

### * NOTE: N = 3 for Runs 3 and 6 for parental x offspring pCO2 treamtents 

```{r about data}

# number of runs 
unique(df$Expeirment) # 1,2,3,4,6 - what are these?
unique(df$Generation) # 1, 2, 3 - generations for larval exposures 

# Lets just see our replication strength per treatment in this summary file..
Rep_Summary_all <- df %>% dplyr::select(c(Generation, Year,  Expeirment, Parent_pCO2, Larvae_pCO2)) %>% 
                           dplyr::group_by_all() %>% 
                           dplyr::summarise(n=n())

unique(Rep_Summary_all$n) # we see 3-5 reps per treatment, veery niiiice!

# print the runs
# summary: 
# Expeirment 6 == Generation 3  full cross parent x offspring exposure *** TARGET FOR THIS PAPER! 

Rep_Summary_E6 <- Rep_Summary_all %>%  dplyr::filter(Expeirment %in% 6) # full cross design - match parent x pffspring contributed to F3s juvenile grow out
Rep_Summary_E6 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```

# Divide and conquer - we have 6 different experimens as follows:

```{r what expeirments do we have data for?}
unique(df$Expeirment)
# 1 2 3 4 6
```
    
  
```{r call target datasets}

df_Exp6 <- df %>% filter(Expeirment=="6") # F3 - full factorial
```

# Plot data as barplots mean +- SE for each run

## Manuscript 4; Expeirment 6

**output folder** = '4_F3_Full_factorial'

```{r Mean table + plots; F3 Experiment #6}

# Summarise Percent Deformities for plotting 
Exp6_mean_deformaties <- df_Exp6 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Percent.deformed) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                           n           = n(),
                                           sd_Perc_def   = sd(Percent.deformed),
                                           se_Perc_def   = sd_Perc_def/(sqrt(n)))

# Barplots
Exp6_barplot <- ggplot(Exp6_mean_deformaties) +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_def-se_Perc_def, 
                                           ymax=mean_Perc_def+se_Perc_def), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_def,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="Experiment #6 (F3)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                        scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

print(Exp6_barplot)


# Boxplot
Exp6_boxplot <- ggplot(df_Exp6,
                       aes(x=Larvae_pCO2, 
                                         y=Percent.deformed,
                                         fill=factor(Larvae_pCO2)), 
                                  stat="identity") +
                        geom_boxplot() +
                        geom_jitter(width = 0.25) +
                        labs(title="Experiment #6 (F3)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                        scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=10),
                              legend.position="none")   +
                        facet_wrap(~Parent_pCO2)
print(Exp6_boxplot)

# Mean SE
Exp6_MeanSE <- ggplot(df_Exp6,aes(x=Larvae_pCO2, 
                                 y=Percent.deformed, 
                                 colour=Larvae_pCO2)) +
                             # scale_linetype(c("dotted","solid")) +
                        scale_colour_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                        stat_summary(fun.y="mean", size = 0.8, color = "black",
                                          position = position_dodge2(width = 1)) +
                        stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 1)) +
                        labs(title="Experiment #6 (F3)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              # axis.title.y=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              plot.title = element_text(size=12),
                              legend.position="none")  +
                        facet_wrap(~Parent_pCO2)
print(Exp6_MeanSE)


# output yo stuff homie!
write.csv(Exp6_mean_deformaties,"Output/Physiology/Dstage_abnormalities/Mean_Abnormalities.csv")

pdf("Output/Physiology/Dstage_abnormalities/MeanSE_Abnormalities.pdf", width = 5, height = 12) # three figures
ggarrange(
  Exp6_barplot,
  Exp6_boxplot,
  Exp6_MeanSE,
  nrow = 3
)
dev.off()
```



# Statistics 

## Manuscript 4; Expeirment 6

**output folder** = '4_F3_Full_factorial'

**about**: two way analysis only larval treatment

```{r  Statistics; F2 Experiment #3} 

aovMOD_Exp6 <- lm(Percent.deformed ~ Larvae_pCO2 * Parent_pCO2, data=df_Exp6)
shapiro.test(resid(aovMOD_Exp6)) # 0.8472 - norm
leveneTest(aovMOD_Exp6) # 0.4374- pass
summary(aov(aovMOD_Exp6))

#                         Df  Sum Sq Mean Sq F value  Pr(>F)   
# Larvae_pCO2              2 0.17830 0.08915   6.388 0.00853 **
# Parent_pCO2              2 0.01225 0.00612   0.439 0.65193   
# Larvae_pCO2:Parent_pCO2  4 0.12531 0.03133   2.245 0.10708   
# Residuals               17 0.23724 0.01396

capture.output(summary(aov(aovMOD_Exp6)),file="Output/Physiology/Dstage_abnormalities/ANOVA.doc") 


aovMOD_Exp6_means_contr <- emmeans::emmeans(object = aovMOD_Exp6, # run tukey
                                    pairwise ~ "Larvae_pCO2",
                                    adjust = "tukey")
aovMOD_Exp6_mod_means   <- multcomp::cld(object = aovMOD_Exp6_means_contr$emmeans, Letters = letters) # letter display\

 # Larvae_pCO2   emmean     SE df lower.CL upper.CL .group
 # Low pCO2       0.312 0.0394 17    0.229    0.395  a    
 # Moderate pCO2  0.445 0.0394 17    0.362    0.528  ab   
 # High pCO2      0.522 0.0425 17    0.432    0.612   b

capture.output(aovMOD_Exp6_mod_means,file="Output/Physiology/Dstage_abnormalities/TukeyHSD.doc") 


# plot the associated effect just the pCO2 exposure treatment
pCO2exposureEffect <- ggplot(df_Exp6,aes(x=Larvae_pCO2, 
                                 y=Percent.deformed, 
                                 colour=Larvae_pCO2)) +
                             # scale_linetype(c("dotted","solid")) +
                        scale_colour_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                        stat_summary(fun.y="mean", size = 0.8, color = "black",
                                          position = position_dodge2(width = 1)) +
                        stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 1)) +
                        labs(title="Experiment #6 (F3)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              # axis.title.y=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              plot.title = element_text(size=12),
                              legend.position="none")


pdf("Output/Physiology/Dstage_abnormalities/MeanSE_Abnormalities_pCO2ExposureEffect.pdf", width = 5, height = 4) # three figures
print(
 pCO2exposureEffect
)
dev.off()
```

