---
title: "Survival D stage to metammorphosis"
author: "Sam Gurr"
date: "9/29/2023"
output:
  pdf_document: default
  html_document: default
---


# Survival D stage to metammorphosis

**About**: These data presented here are relative to the D-stage density within tank,
useful comparison to the data from stocking densitry (as embryos) to metamorphosis.
Possile benefit of this analysis is the cross-generation comparison, as fertilization success differed
strongly between treatment and generation thus the weight of treatemnt and/or generation effect could be skewed toward 
the inital development. Instead, here the survival analysis is investigated after D-hinge development giving a 
perspective of treatment/generation effect across the larval period. 

### SET UP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_multigen_OA/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's work
```

### LOAD PACKAGES
```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(survival)
library(Rmisc)
library(coxme)
library(survminer)
library(ggsurvfit) # survfit2
library(gtsummary) # tbl_survfit

```

### LOAD DATA
```{r, echo=FALSE}

# note: we have percent_survival_rel_embryo and Percent_Survival_rel_Dstage 
# experiment #4 does not have percent relative to embryo, but is percent rel to d hinge stage so we will use that 
# the survival at Dstage as abnormality data can be used for the inital assessment 

df<- read.csv("Data/larval_survival.csv", header = T) %>% 
                  dplyr::mutate(Percent_Survival = as.numeric(Percent_Survival_rel_dstage,
                                Percent_Survival.embryo = as.numeric(Percent_Survival_rel_stocking))) %>% 
                  # dplyr::select(-treatment) %>%  # redundant column with Exposure_OA
                  
                  # rename the variables for consistency using 'case_when'
                  dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                                        Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                                        Parental_OA ==  "HIGH" ~ "High pCO2",
                                                        Parental_OA ==  "WILD" ~ "None")) %>% 
                  dplyr::mutate(Exposure_OA = case_when(Exposure_OA ==  "LOW" ~ "Low pCO2", 
                                                        Exposure_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                                        Exposure_OA ==  "HIGH" ~ "High pCO2")) %>% 
                  
                  # change the levels for order when plotting - call as factors too!
                  dplyr::mutate(Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, 
                                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
                  dplyr::mutate(Exposure_OA = as.factor(forcats::fct_relevel(Exposure_OA, 
                                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
                  
                  # rename them
                  dplyr::rename(Parent_pCO2 = Parental_OA) %>% 
                  dplyr::rename(Larvae_pCO2  = Exposure_OA) # done 
  
  
df$Percent_Survival_trans <- asin(sqrt(df$Percent_Survival))

# head(df)
# str(df)
```

# About data 

### Experiments: 

  - (2) F1 larvae under three OA levels, *grown out to adult stage*

  - (4) F2 larvae under two OA levels, *grown out to adult stage*

  - (6) F3 larvae under 3 x 3 parent x offsring OA, *no grow out past juvenile takedown of the experiment*

### * NOTE: N = 3 for Runs 3 and 6 for parental x offspring pCO2 treamtents 

```{r about data}

# number of runs 
unique(df$Generation) # 1, 2, and 3

Rep_Summary_all <- df %>% dplyr::select(c(Generation, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_G1 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 1) # 
Rep_Summary_G1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G2 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 2) #
Rep_Summary_G2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G3 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 3) # 
Rep_Summary_G3 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```


### Build data frames for each expeirment 


```{r}
df_Exp2 <- df%>%
  filter(Generation=="1") %>% dplyr::mutate(Experiment = 2)

df_Exp4 <- df%>%
  filter(Generation=="2") %>% dplyr::mutate(Experiment = 4)

df_Exp6 <- df%>%
  filter(Generation=="3") %>% dplyr::mutate(Experiment = 6)
```


# Plot data as barplots mean +- SE for each run

## Manuscript 1; Expeirments 2 and 4

**output folder** = '1_Survival_Growth'

* note that metamorphosis was observed at 10 dpf in F1s and 14 dpf in F2s, therfore 
larval survival comapred across gnerations should only show < 10 dpf survival 
 
```{r Mean table + plots; F1 Experiment #2}

stL_Exp2 <- summarySE(df_Exp2, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_Exp2 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


# plot relative to stocking density aas embryos 



# plot relative to 100$% restart at D stage 
Exp2_PlotMeanSE <-  ggplot(stL_Exp2, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F1 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
Exp2_PlotMeanSE

Exp2_larvae_PlotMeanSE <-  stL_Exp2 %>%  dplyr::filter(Age <= 10) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F1 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
Exp2_larvae_PlotMeanSE

# output the data !!

write.csv(stL_Exp2,"Output/1_Survival_Growth/Experiment1_Mean_Survival.csv")

pdf("Output/1_Survival_Growth/Experiment1_F1larvae_Survival_MeanSE.pdf", width=10, height=5)
print(Exp2_larvae_PlotMeanSE)
dev.off()

pdf("Output/1_Survival_Growth/Experiment1_F1_Survival_MeanSE.pdf", width=10, height=5)
print(Exp2_PlotMeanSE)
dev.off()
```

```{r Mean table + plots; F2 Experiment #4}

stL_Exp4 <- summarySE(df_Exp4, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_Exp4 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Exp4_PlotMeanSE <-  ggplot(stL_Exp4, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F2 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
Exp4_PlotMeanSE

Exp4_larvae_PlotMeanSE <-  stL_Exp4 %>%  dplyr::filter(Age <= 14) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F2 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
Exp4_larvae_PlotMeanSE

# output the data !!

write.csv(stL_Exp4,"Output/1_Survival_Growth/Experiment4_Mean_Survival.csv")

pdf("Output/1_Survival_Growth/Experiment4_F2larvae_Survival_MeanSE.pdf", width=10, height=5)
print(Exp4_larvae_PlotMeanSE)
dev.off()

pdf("Output/1_Survival_Growth/Experiment4_F2_Survival_MeanSE.pdf", width=10, height=5)
print(Exp4_PlotMeanSE)
dev.off()
```

## Manuscript 4; Expeirments 6

**output folder** = '4_F3_Full_factorial'

```{r Mean table + plots; F3 Experiment #6}

stL_Exp6 <- summarySE(df_Exp6, measurevar="Percent_Survival", groupvars=c("Age", "Parent_pCO2", "Larvae_pCO2"))
stL_Exp6 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Exp6_PlotMeanSE <-  ggplot(stL_Exp6, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F3 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black")) + 
                          facet_wrap(~Parent_pCO2)
Exp6_PlotMeanSE

Exp6_larvae_PlotMeanSE <-  stL_Exp6 %>%  dplyr::filter(Age <= 12) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F3 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black")) + 
                          facet_wrap(~Parent_pCO2)
 
Exp6_larvae_PlotMeanSE

# output the data !!

write.csv(stL_Exp6,"Output/4_F3_Full_factorial/survival/Experiment6_Mean_Survival.csv")

pdf("Output/4_F3_Full_factorial/survival/Experiment6_F3_larvae_Survival_MeanSE.pdf", width=10, height=5)
print(Exp6_larvae_PlotMeanSE)
dev.off()

```

# Kaplan-Meier plots


## About:

* We will use the {ggsurvfit} package to generate Kaplan-Meier plots. This package aims to ease plotting of time-to-event endpoints using the power of the {ggplot2} package. See http://www.danieldsjoberg.com/ggsurvfit/index.html for details.

## Build 'surv' object for statistical analysis 

* useful links [https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Comparing_survival_times_between_groups]

* What the hell does a surv object mean?? 
  
  - a 'surv' object reads ALL events as that of a dead (1) or alive (0) individual
  
  - but but but... we have volumetric count data of live and dead! Does this not throw a wrench into this type of analysis? 
  
    - Nope! Below we use logic to create a GIGANTIC table formatted correctly for this purpose!

## Format

 * Build new column for count dead - take the count alive data from time 0 and porp survival to calcualted number dead

## Convert to binary formant (super long data tbale!) using two for loops:

 * (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting
 
 * (2) Mean counts alive and dead by tank - better represented for plots!






## Manuscript 1; Expeirments 2 and 4

**output folder** = '1_Survival_Growth'

**note**: out ultimate objsetice with the sruvival data in this manuscript is to investigate change insurvivroship in reponse to treamtent across treatment 
and generational timescales - we reached metamorphosis at 10 DPF for the Experiment 2 and at 14 DPF or experiment 4, so we will need 
to cater to this difference to make meaningful comparisons





### Experiment 2 F1s


```{r format F1 Experiment #2}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_Exp2 <- df_Exp2 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(proportion = Percent_Survival/100,
                        Count_total = case_when(proportion == 0 ~ 0,
                                                proportion >0 ~
                                                 (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive)),# proportion is that of alive
                           Count_dead = Count_total - Count_alive)
```

```{r convert to binary F1 Experiment #2}

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
binary_Exp2_all     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_Exp2)) {
   #count_alive <- round(df_Exp2[i,]$Count_alive/100) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_Exp2[i,]$Count_dead/100)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_Exp2[i,]$Count_alive) # raw counts
   count_dead  <- round(df_Exp2[i,]$Count_dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_Exp2[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_Exp2[i,]$Larvae_pCO2,
                            Generation  = df_Exp2[i,]$Generation,
                            Tank        = df_Exp2[i,]$Tank,  
                            Age         = df_Exp2[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp2_all <- rbind(binary_Exp2_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_Exp2_all$Tank) # "A" "B" "C" "D" "E" NA
binary_Exp2_all <- binary_Exp2_all %>% 
                    dplyr::mutate(Generation = 1,
                                  Tank = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3,
                                                   Tank== 'D'~4,
                                                   Tank== 'E'~5)) %>% 
                    na.omit()


# (2) Mean counts alive and dead by tank - better representated for plots!
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_Exp2_MEANS <- df_Exp2 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))

# call datafarme binary for plotting
binary_Exp2_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_Exp2_MEANS)) {
   #count_alive <- round(df_Exp2_MEANS[i,]$mean_Count_alive/100) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_Exp2_MEANS[i,]$mean_Count_dead/100)  # divide all counts by 1000, round to nearest

   count_alive <- round(df_Exp2_MEANS[i,]$mean_Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp2_MEANS[i,]$mean_Count_dead)  # divide all counts by 1000, round to nearest
   
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_Exp2_MEANS[i,]$Larvae_pCO2,
                            Age         = df_Exp2_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp2_MEANS <- rbind(binary_Exp2_MEANS,loopDF) #bind to a cumulative list dataframe
  
}
binary_Exp2_MEANS <- binary_Exp2_MEANS %>% 
                      dplyr::mutate(Generation = 1) %>% 
                      na.omit()

# test whether 0s and 1s do pan out to the count dead and alive
# length((binary_Exp2_all %>% dplyr::filter(c(Tank %in% '1' & Age == 4 & Larvae_pCO2 %in% 'Low pCO2' & Count_dead == 1)))$Count_dead)
# length((binary_Exp2_all %>% dplyr::filter(c(Tank %in% '1' & Age == 4 & Larvae_pCO2 %in% 'Low pCO2' & Count_dead == 0)))$Count_dead)
# # 58667 # the coutn alive 
# #32583 # eaxactly right, we counted dead == 1 and alive == 0

```

**note**: larval period ended at first sign of metamorphossi at 10 DPF  - truncate before this age to compare
```{r filter larval period F1 Experiment #2}

binary_Exp2_larvae      <- binary_Exp2_all %>% filter(Age <=10) %>% # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
                                        dplyr::mutate(Prop_time_metamophosis = Age/10)
binary_Exp2_larvaeMEANS <- binary_Exp2_MEANS %>% filter(Age <=10)  %>% # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
                                        dplyr::mutate(Prop_time_metamophosis = Age/10)

```

```{r surv object F1 Experiment #2}

survExp2_all <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)


# convert Age to the proportion of time to metamorphosis 
# binary_Exp2_larvaeMEANS$Prop_time_metamophosis <- (binary_Exp2_larvaeMEANS$Age / 10)


survExp2          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)
survExp2_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)

?survfit2
```

```{r plot Kaplan Meier F1 Experiment #2}

Exp2_larvaeKaplanMeier.age <- survExp2 %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1,type ="survival") +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

# plot the surv data
pdf("Output/1_Survival_Growth/Expeirment2_Survival_KaplanMeier_age.pdf", width=5, height = 4)
Exp2_larvaeKaplanMeier.age
dev.off()

Exp2_larvaeKaplanMeier.prop <- survExp2_proptime %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

# plot the surv data
pdf("Output/1_Survival_Growth/Expeirment2_Survival_KaplanMeier_proptime.pdf", width=5, height = 4)
Exp2_larvaeKaplanMeier.prop
dev.off()

Exp_10d_surv_prob <- survExp2 %>% 
                          tbl_survfit(
                            times = 10,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )


```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).
* We get the log-rank p-value using the survdiff function
```{r log-rank test survdiff F1 Experiment #2}

Exp2_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)
#                               N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      11541     9682     9688   0.00342    0.0103
# Larvae_pCO2=Moderate pCO2  9424     8038     8531  28.47029   82.1467
# Larvae_pCO2=High pCO2      9564     9136     8637  28.77948   83.4968

```

# The Cox regression model
* We may want to quantify an effect size for a single variable, or include **more than one variable** into a regression model to account for the effects of multiple variables.
* The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.
```{r cox proprtional hazards F1 Experiment #2}
library(survminer) # ggforect hazard ratio plot

#  Experiment 2

# Cox proportional hazards regression model
Exp2_coxph_test          <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                 data = binary_Exp2_larvaeMEANS)
Exp2_hazardratio.coxph   <- ggforest(Exp2_coxph_test, data = binary_Exp2_larvaeMEANS)

Exp2_coxph_test_table   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                 data = binary_Exp2_larvae) %>%  
                             tbl_regression(exp = TRUE) 


Exp2_coxph_test.frailty         <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2 + frailty(ID), 
                                         data = binary_Exp2_larvae) 
Exp2_hazardratio.coxp.frailty   <- ggforest(Exp2_coxph_test.frailty, data = binary_Exp2_larvae)




# mixed Cox proportional hazards regression model
binary_Exp2_larvae$ID <- factor(paste0(
                              substr(binary_Exp2_larvae$Larvae_pCO2,1,1),
                              binary_Exp2_larvae$Tank))
binary_Exp2_larvae    <- droplevels(binary_Exp2_larvae)

Exp2_coxme_test         <- coxme(formula = Surv(Age, Count_dead) ~ Larvae_pCO2 +
                                   (1 | ID), data = binary_Exp2_larvae)
Exp2_coxme_test_table   <- coxme(Surv(Age, Count_dead) ~ Larvae_pCO2 + 
                                   (1 | ID), data = binary_Exp2_larvae) %>%  
                             tbl_regression(exp = TRUE) 
# * NOTE cannot make a hazard ratio plot with coxme (not compatible with ggforest) therefore adjust based on this output

print(Exp2_coxme_test_table)
# output yo stuff homie!
capture.output(Exp2_coxme_test,
               file="Output/1_Survival_Growth/survival/Experiment2_Survival_CoxReg.doc")
capture.output(print(Exp2_coxme_test_table),
               file="Output/1_Survival_Growth/Experiment2_Survival_CoxReg_HR.txt")

pdf("Output/1_Survival_Growth/survival/Experiment2_HazardRatio.pdf", width = 6, height = 6) # three figures
Exp2_hazardratio.coxph
dev.off()

```






### Experiment 4 F2s

```{r format F2 Experiment #4}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_Exp4 <- df_Exp4 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(proportion = Percent_Survival/100,
                        Count_total = case_when(proportion == 0 ~ 0,
                                                proportion >0 ~
                                                 (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive)),# proportion is that of alive
                        Count_dead = case_when(proportion == 0 ~ 0,
                                                proportion >0 ~ Count_total - Count_alive))
```

```{r convert to binary F2 Experiment #4}

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
binary_Exp4_all     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_Exp4)) {
   #count_alive <- round(df_Exp4[i,]$Count_alive/100) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_Exp4[i,]$Count_dead/100)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_Exp4[i,]$Count_alive) # raw counts
   count_dead  <- round(df_Exp4[i,]$Count_dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_Exp4[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_Exp4[i,]$Larvae_pCO2,
                            Generation  = df_Exp4[i,]$Generation,
                            Tank        = df_Exp4[i,]$Tank,  
                            Age         = df_Exp4[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp4_all <- rbind(binary_Exp4_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_Exp4_all$Tank) # "A" "B" "C" "D" "E" NA
binary_Exp4_all <- binary_Exp4_all %>% 
                    dplyr::mutate(Generation = 2,
                                  Tank_num = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3)) %>% 
                    dplyr::mutate(Tank = paste0(substr(Parent_pCO2,1,1),
                                                substr(Larvae_pCO2,1,1),
                                                '_',
                                                Tank_num)) %>% 
                    na.omit()


# (2) Mean counts alive and dead by tank - better representated for plots!
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_Exp4_MEANS <- df_Exp4 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))

# call datafarme binary for plotting
binary_Exp4_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_Exp4_MEANS)) {
   #count_alive <- round(df_Exp4_MEANS[i,]$mean_Count_alive/100) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_Exp4_MEANS[i,]$mean_Count_dead/100)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_Exp4_MEANS[i,]$mean_Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp4_MEANS[i,]$mean_Count_dead)  # divide all counts by 1000, round to nearest
   
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_Exp4_MEANS[i,]$Larvae_pCO2,
                            Age         = df_Exp4_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp4_MEANS <- rbind(binary_Exp4_MEANS,loopDF) #bind to a cumulative list dataframe
  
}
binary_Exp4_MEANS <- binary_Exp4_MEANS %>% 
                      dplyr::mutate(Generation = 2) %>% 
                      na.omit()

```

**note**: larval period ended at first sign of metamorphosis at 14 DPF  - truncate before this age to compare
```{r filter larval period F2 Experiment #4}

binary_Exp4_larvae      <- binary_Exp4_all %>% filter(Age <=14) %>% # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
                            dplyr::mutate(Prop_time_metamophosis = Age/14)
binary_Exp4_larvaeMEANS <- binary_Exp4_MEANS %>% filter(Age <=14) %>% # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
                            dplyr::mutate(Prop_time_metamophosis = Age/14)
```

```{r surv object F2 Experiment #4}

survExp4_all <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)

# convert Age to the proportion of time to metamorphosis 
binary_Exp4_larvaeMEANS$Prop_time_metamophosis <- (binary_Exp4_larvaeMEANS$Age / 14)

survExp4          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
survExp4_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
```

```{r plot Kaplan Meier F2 Experiment #4}

Exp4_larvaeKaplanMeier.age <- survExp4 %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:14)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

# plot the surv data
pdf("Output/1_Survival_Growth/survival/Expeirment4_Survival_KaplanMeier.age.pdf", width=5, height = 4)
Exp4_larvaeKaplanMeier.age
dev.off()

Exp4_larvaeKaplanMeier.prop <- survExp4_proptime %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)
# plot the surv data
pdf("Output/1_Survival_Growth/survival/Expeirment4_Survival_KaplanMeier.prop.pdf", width=5, height = 4)
Exp4_larvaeKaplanMeier.prop
dev.off()

Exp4_14d_surv_prob <- survExp4 %>% 
                          tbl_survfit(
                            times = 14,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )
```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).
* We get the log-rank p-value using the survdiff function
```{r log-rank test survdiff F2 Experiment #4}

Exp4_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
#                             N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      250      166      152     1.200      3.84
# Larvae_pCO2=Moderate pCO2 336      190      204     0.899      3.84

```

# The Cox regression model
* We may want to quantify an effect size for a single variable, or include **more than one variable** into a regression model to account for the effects of multiple variables.
* The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.
```{r cox & hazard ratio F2 Experiment #4}
library(survminer) # ggforect hazard ratio plot


binary_Exp4_larvae$Larvae_pCO2 <- factor(binary_Exp4_larvae$Larvae_pCO2, exclude = 'High pCO2')

#  Experiment 4

# Cox proportional hazards regression model
Exp4_coxph_test         <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvae) 
Exp4_hazardratio_plot   <- ggforest(Exp4_coxph_test, data = binary_Exp4_larvae)
Exp4_coxph_test_table   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvae) %>%  
                             tbl_regression(exp = TRUE) 

# mixed Cox proportional hazards regression model
binary_Exp4_larvae$ID <- factor(paste0(
                              substr(binary_Exp4_larvae$Larvae_pCO2,1,1),
                              binary_Exp4_larvae$Tank))
binary_Exp4_larvae <- droplevels(binary_Exp4_larvae)

Exp4_coxme_test         <- coxme(formula = Surv(Age, Count_dead) ~ Larvae_pCO2 +
                                   (1 | ID), data = binary_Exp4_larvae)
Exp4_coxme_test_table   <- coxme(Surv(Age, Count_dead) ~ Larvae_pCO2 + 
                                   (1 | ID), data = binary_Exp4_larvae) %>%  
                             tbl_regression(exp = TRUE) 
# * NOTE cannot make a hazard ratio plot with coxme (not compatible with ggforest) therefore adjust based on this output


# output yo stuff homie!
capture.output(Exp4_coxph_test,file="Output/1_Survival_Growth/survival/Experiment4_Survival_CoxReg.doc")
capture.output(Exp4_coxme_test_table,file="Output/1_Survival_Growth/survival/Experiment4_Survival_CoxReg_HR.doc")


pdf("Output/1_Survival_Growth/survival/Experiment4_HazardRatio.pdf", width = 10, height = 6) # three figures
Exp4_hazardratio_plot
dev.off()


```




### Experiment 2 and 4 MERGE DATA

**about**: we have parallel datasets to use in a signle manuscript - merge the binary files and run stats / plots

**note**: the proportion time to metamorphosis column is *already accounted for* with the different times to metamorphossis!

* NOTE: omit the 'high' treatment in F1 as this was not present in F2 larvae
```{r merge Experiments 2 and 4}
# all - with tank replicate included

binary_Exp2_Exp4_all             <- rbind( (binary_Exp2_larvae),# %>% dplyr::select(!cluster)),
                                           (binary_Exp4_larvae)# %>% dplyr::select(!Tank_num))
                                           ) %>% 
                                         dplyr::filter(!(Larvae_pCO2 == 'High pCO2')) %>% 
                                         dplyr::mutate(Prop_time_metamophosis = 
                                                            case_when(Generation == 1 ~ (Age / 10),
                                                                      Generation == 2 ~ (Age / 14))) %>% 
                                         # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
                                         dplyr::filter(!Prop_time_metamophosis >1) 

# omit high pCO2 as a level in larvae pCO2 - just compare moderate and low here
binary_Exp2_Exp4_all$Larvae_pCO2 <- factor(binary_Exp2_Exp4_all$Larvae_pCO2, exclude = 'High pCO2')

#  by mean tank
binary_Exp2_Exp4_MEANS             <- rbind(binary_Exp2_larvaeMEANS, binary_Exp4_larvaeMEANS ) %>% 
                                           dplyr::filter(!(Larvae_pCO2 == 'High pCO2')) %>% 
                                           dplyr::mutate(Prop_time_metamophosis = 
                                                            case_when(Generation == 1 ~ (Age / 10),
                                                                      Generation == 2 ~ (Age / 14))) %>% 
                                         # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
                                         dplyr::filter(!Prop_time_metamophosis >1) 
binary_Exp2_Exp4_MEANS$Larvae_pCO2 <- factor(binary_Exp2_Exp4_MEANS$Larvae_pCO2, exclude = 'High pCO2')


# sanity check
binary_Exp2_Exp4_all %>% dplyr::filter(Prop_time_metamophosis > 1) # should be no data!
binary_Exp2_Exp4_MEANS %>% dplyr::filter(Prop_time_metamophosis > 1) # should be no data!


# Subsets to explore further 
binary_Exp2_Exp4_larvaeLOW       <- binary_Exp2_Exp4_all %>% filter(Larvae_pCO2 %in% "Low pCO2")
binary_Exp2_Exp4_MEANS_larvaeLOW <- binary_Exp2_Exp4_MEANS %>% filter(Larvae_pCO2 %in% "Low pCO2") 

binary_Exp2_Exp4_larvaeMOD       <- binary_Exp2_Exp4_all %>% filter(Larvae_pCO2 %in% "Moderate pCO2")
binary_Exp2_Exp4_MEANS_larvaeMOD <- binary_Exp2_Exp4_MEANS %>% filter(Larvae_pCO2 %in% "Moderate pCO2") 

# what days are included?
unique(binary_Exp2_larvae$Age) # 2  4  8 10
unique(binary_Exp4_larvae$Age) # 2  6  8 10
unique(binary_Exp2_Exp4_all$Age) #  2  4  8 10  6 14

unique(binary_Exp2_larvaeMEANS$Age) # 2  4  8 10
unique(binary_Exp4_larvaeMEANS$Age) # 2  6  8 10
unique(binary_Exp2_Exp4_MEANS$Age) #  2  4  8 10  6 14

```

## Surv object with all data v. mean data
* note you can see the CI for the 'all' data surv object is small - these are 0s and 1s for each
* note there is a large CI for the 'mean' data 
```{r surv object merge F1 & F2 for plotting}

s1_2          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2+Generation, data = binary_Exp2_Exp4_MEANS)
s1_2_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2+Generation, data = binary_Exp2_Exp4_MEANS)


s1_2LOW          <- survfit2(Surv(Age, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeLOW)
s1_2proptimeLOW  <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeLOW)

s1_2MOD          <- survfit2(Surv(Age, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeMOD)
s1_2proptimeMOD  <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeMOD)
```

```{r plot Kaplan Meier merge F1 & F2}

Exp2_Exp4_larvaeKaplanMeier <- s1_2_proptime %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("green","forestgreen",
                                                      "orange","darkorange4")) +
                        scale_fill_manual(values = c("green","forestgreen",
                                                     "orange","darkorange4")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)

# output plot 
pdf(paste0("Output/1_Survival_Growth/survival/Experiments2&4_KaplanMeier.pdf"),width=8, height=4)
print(Exp2_Exp4_larvaeKaplanMeier)
dev.off()


Exp2_Exp4_larvaeKaplanMeierLOW <- s1_2proptimeLOW %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("green","forestgreen")) +
                        scale_fill_manual(values = c("green","forestgreen")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)
# output plot 
pdf(paste0("Output/1_Survival_Growth/survival/Experiments2&4_KaplanMeier_LowpCO2.pdf"),width=8, height=4)
print(Exp2_Exp4_larvaeKaplanMeierLOW)
dev.off()



Exp2_Exp4_larvaeKaplanMeierMOD <- s1_2proptimeMOD %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("orange","darkorange")) +
                        scale_fill_manual(values = c("orange","darkorange")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)
# output plot 
pdf(paste0("Output/1_Survival_Growth/survival/Experiments2&4_KaplanMeier_ModeratepCO2.pdf"),width=8, height=4)
print(Exp2_Exp4_larvaeKaplanMeierMOD)
dev.off()


```

```{r log-rank test survdiff  merge F1 & F2}

F1_F2_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+Generation, data = binary_Exp2_Exp4_MEANS)
#                                             N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2, Generation=1      11541     9682     9230     22.18     85.47
# Larvae_pCO2=Low pCO2, Generation=2        250      166      310     66.67    161.11
# Larvae_pCO2=Moderate pCO2, Generation=1  9424     8038     8122      0.87      2.98
# Larvae_pCO2=Moderate pCO2, Generation=2   336      190      415    121.78    296.27

```

```{r cox & hazard ratio merge F1 & F2}

binary_Exp2_Exp4_all$Generation <- as.character(binary_Exp2_Exp4_all$Generation)
binary_Exp2_Exp4_all$Gen_pCO2   <- paste0(as.character(binary_Exp2_Exp4_all$Generation),
                                          '_',
                                          (binary_Exp2_Exp4_all$Larvae_pCO2))
binary_Exp2_Exp4_all$ID         <- paste0(binary_Exp2_Exp4_all$Tank, # true random factor for the coxme model
                                          '_',
                                          binary_Exp2_Exp4_all$Gen_pCO2)


#  Experiments 2 and 4 compare common treatments with surv at proportion time to metamorphosis

# Cox proportional hazards regression model
Exp2_Exp4_coxph_test.int  <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) 
                                    ~ Gen_pCO2, data = binary_Exp2_Exp4_all)
Exp2_Exp4_hazardratio.int <- ggforest(Exp2_Exp4_coxph_test.int, data = binary_Exp2_Exp4_all)
Exp2_Exp4_coxph_table.int <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ Gen_pCO2, 
                                    data = binary_Exp2_Exp4_all) %>%  
                                      tbl_regression(exp = TRUE) 


Exp2_Exp4_coxph_test.add  <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) 
                                    ~ Generation + Larvae_pCO2, data = binary_Exp2_Exp4_all)
Exp2_Exp4_hazardratio.add <- ggforest(Exp2_Exp4_coxph_test.add, data = binary_Exp2_Exp4_all)
Exp2_Exp4_coxph_table.add <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ 
                                     Generation + Larvae_pCO2, 
                                    data = binary_Exp2_Exp4_all) %>%  
                                      tbl_regression(exp = TRUE) 

pdf("Output/1_Survival_Growth/Experiments2&4_HazardRatio.pdf", width = 10, height = 6) # three figures
ggarrange(Exp2_Exp4_hazardratio.int, Exp2_Exp4_hazardratio.add, ncol=2)
dev.off()

# MIXED COX PROPRTION HAZARDS MODEL 

# interaction
binary_Exp2_Exp4_all$ID <- factor(binary_Exp2_Exp4_all$ID)
binary_Exp2_Exp4_all     <- droplevels(binary_Exp2_Exp4_all)
Exp2_Exp4_coxme_test         <- coxme(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Gen_pCO2 +
                                   (1 | ID), data = binary_Exp2_Exp4_all)
Exp2_Exp4_coxme_test_table   <- coxme(Surv(Prop_time_metamophosis, Count_dead) ~ Gen_pCO2 + 
                                   (1 | ID), data = binary_Exp2_Exp4_all) %>%  
                             tbl_regression(exp = TRUE) 

#additive mixed
Exp2_Exp4_coxme_all.add         <- coxme(formula = Surv(Prop_time_metamophosis, Count_dead) ~ 
                                           Generation + Larvae_pCO2 + (1 | ID), 
                                         data = binary_Exp2_Exp4_all)
Exp2_Exp4_coxme_all_table.add  <- coxme(formula = Surv(Prop_time_metamophosis, Count_dead) ~ 
                                           Generation + Larvae_pCO2 + (1 | ID), 
                                         data = binary_Exp2_Exp4_all) %>%  
                                        tbl_regression(exp = TRUE) 

capture.output(Exp2_Exp4_coxme_test,file="Output/1_Survival_Growth/survival/Experiments2&4_Survival_CoxReg_int.doc")
capture.output(Exp2_Exp4_coxme_all.add,file="Output/1_Survival_Growth/survival/Experiments2&4_Survival_CoxReg_add.doc")
capture.output(Exp2_Exp4_coxme_test_table,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxReg_HR.doc")



#  Experiments 2 and 4 compare Low pCO2

binary_Exp2_Exp4_larvaeLOW$Generation <- as.character(binary_Exp2_Exp4_larvaeLOW$Generation)
binary_Exp2_Exp4_larvaeLOW$ID         <- paste0(binary_Exp2_Exp4_larvaeLOW$Tank, # true random factor for the coxme model
                                          '_',
                                          binary_Exp2_Exp4_larvaeLOW$Generation)

# Cox proportional hazards regression model
Exp2_Exp4_coxph_testLOW       <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_larvaeLOW)
Exp2_Exp4_hazardratio_plotLOW <- ggforest(Exp2_Exp4_coxph_testLOW, 
                                          data = binary_Exp2_Exp4_larvaeLOW)
Exp2_Exp4_coxph_test_tableLOW <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_larvaeLOW) %>% tbl_regression(exp = TRUE) 

# mixed Cox proportional hazards regression model
binary_Exp2_Exp4_larvaeLOW$ID   <- factor(binary_Exp2_Exp4_larvaeLOW$ID )
binary_Exp2_Exp4_larvaeLOW      <- droplevels(binary_Exp2_Exp4_larvaeLOW)

Exp2_Exp4_coxme_testLOW         <- coxme(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation +
                                   (1 | ID), data = binary_Exp2_Exp4_larvaeLOW)
Exp2_Exp4_coxme_test_tableLOW   <- coxme(Surv(Prop_time_metamophosis, Count_dead) ~ Generation + 
                                   (1 | ID), data = binary_Exp2_Exp4_larvaeLOW) %>%  
                                        tbl_regression(exp = TRUE) 
?coxme

capture.output(Exp2_Exp4_coxph_testLOW,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxReg_LowpCO2.doc")
capture.output(Exp2_Exp4_coxme_test_tableLOW,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxReg_LowpCO2_HR.doc")

pdf("Output/1_Survival_Growth/Experiments2&4_HazardRatio_LowpCO2.pdf", width = 10, height = 3) # three figures
print(Exp2_Exp4_hazardratio_plotLOW)
dev.off()





#  Experiments 2 and 4 compare Moderate pCO2

binary_Exp2_Exp4_larvaeMOD$Generation <- as.character(binary_Exp2_Exp4_larvaeMOD$Generation)
binary_Exp2_Exp4_larvaeMOD$ID         <- paste0(binary_Exp2_Exp4_larvaeMOD$Tank, # true random factor for the coxme model
                                          '_',
                                          binary_Exp2_Exp4_larvaeMOD$Generation)
binary_Exp2_Exp4_larvaeMOD$Generation <- as.character(binary_Exp2_Exp4_larvaeMOD$Generation)


# Cox proportional hazards regression model
Exp2_Exp4_coxph_testMOD       <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_larvaeMOD)
Exp2_Exp4_hazardratio_plotMOD <- ggforest(Exp2_Exp4_coxph_testMOD, 
                                          data = binary_Exp2_Exp4_larvaeMOD)
Exp2_Exp4_coxph_test_tableMOD <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_larvaeMOD) %>% tbl_regression(exp = TRUE) 


# mixed Cox proportional hazards regression model
binary_Exp2_Exp4_larvaeMOD$ID   <- factor(binary_Exp2_Exp4_larvaeMOD$ID )
binary_Exp2_Exp4_larvaeMOD      <- droplevels(binary_Exp2_Exp4_larvaeMOD)

#interaction mixed
Exp2_Exp4_coxme_testMOD         <- coxme(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation +
                                   (1 | ID), data = binary_Exp2_Exp4_larvaeMOD)
Exp2_Exp4_coxme_test_tableMOD   <- coxme(Surv(Prop_time_metamophosis, Count_dead) ~ Generation + 
                                   (1 | ID), data = binary_Exp2_Exp4_larvaeMOD) %>%  
                                        tbl_regression(exp = TRUE) 




capture.output(Exp2_Exp4_coxph_testMOD,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxReg_ModeratepCO2.doc")
pdf(file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxReg_ModeratepCO2_HR.pdf")
Exp2_Exp4_coxme_test_tableMOD
dev.off()

pdf("Output/1_Survival_Growth/Experiments2&4_HazardRatio_ModeratepCO2.pdf", width = 10, height = 3) # three figures
print(Exp2_Exp4_hazardratio_plotMOD)
dev.off()
```




### Experiment 2 and 4 PRE AND POST METAMORPHOSIS 

**Objective**: We know that the post-metamorphic oyster dropped off dramatically after 
under the moderate treament (and there werent any under high!) However our reuslts alone for 
ore metamorhpisis (in chunks above) would otherwise present that moderate is beneficial. 
We must present these data post-metamorphosis.

* Below are chunks for the **delta(pre) versus delta(post)** addressed statistically as 
one-way anova models (fixed effect of treatment) separately for pre and post


* Note: we inly need the percent survival data for this, no counts or binary live dead!

* *How does this work since we have offset time to metamorphosis?*
  - Use equal time pre and equal time post to run the models, meaning...
  If F1 metamorphosed at 10 dpf, the percent survival at 6 days (pre) and 14 days (post) 
  represents a comparable dataset to days 10 (pre) and 18 (post) for F2s that metamorphosed at 
  14 dpf. In this example, we investigate the +-4 days frommetamormophosis 
  
```{r Pre v. Post: data set up and sanity checks}

# Experiment 2 (F1s) metamorphosed at 10 DPF
Exp2_PrePost <- df %>% 
                        filter(Generation=="1" & Larvae_pCO2 %in% c('Low pCO2', 'Moderate pCO2')) %>%
                        dplyr::mutate(Experiment = 2)
sort(unique(Exp2_PrePost$Age)) # 2  4  8 10 14 18

# Day 4 == 6 days pre
# Day 18 == 8 days post

Exp2_Pre.calc <- Exp2_PrePost %>% 
                      dplyr::select(c(Age, proportion_rel_dstage, Tank, Larvae_pCO2,Generation)) %>% 
                      dplyr::filter(Age %in% c(4, 10)) %>% 
                      tidyr::pivot_wider(names_from = Age, 
                                         values_from = proportion_rel_dstage,
                                         names_glue = "Age_{.name}") %>% # renmae new cols
                      dplyr::mutate(type       = "pre", # new col fo rpre
                                    delta.perc = Age_10 - Age_4) %>%  # sutract em
                      dplyr::mutate(delta.perc = case_when(delta.perc > 0 ~ 0, # positives make zero
                                                           .default = delta.perc),
                                    abs.delta.perc = abs(delta.perc)) %>% # now take the absolute value
                      na.omit() %>% 
                      dplyr::select(-c(Age_10,Age_4)) # no longer need these, remove to merge with post
  

Exp2_Post.calc <- Exp2_PrePost %>% 
                      dplyr::select(c(Age, proportion_rel_dstage, Tank, Larvae_pCO2,Generation)) %>% 
                      dplyr::filter(Age %in% c(10, 18)) %>% 
                      tidyr::pivot_wider(names_from = Age, 
                                         values_from = proportion_rel_dstage,
                                         names_glue = "Age_{.name}") %>% # renmae new cols
                      dplyr::mutate(type       = "post", # new col fo rpre
                                    delta.perc = Age_18 - Age_10) %>%  # sutract em
                      dplyr::mutate(delta.perc = case_when(delta.perc > 0 ~ 0, # positives make zero
                                                           .default = delta.perc),
                                    abs.delta.perc = abs(delta.perc))%>% # now take the absolute value
                      na.omit() %>% 
                      dplyr::select(-c(Age_18,Age_10)) # no longer need these, remove to merge with post


Exp2_PrePost.calc <- rbind(Exp2_Pre.calc,Exp2_Post.calc)







# Experiment 4 (F2s) metamorphosed at 14 DPF
Exp4_PrePost <- df %>% filter(Generation=="2") %>% dplyr::mutate(Experiment = 4)
sort(unique(Exp4_PrePost$Age)) # 2  6  8 10 14 21

# Day 8  == 6 days pre
# Day 21 == 8 days post

Exp4_Pre.calc <- Exp4_PrePost %>% 
                      dplyr::select(c(Age, proportion_rel_dstage, Tank, Larvae_pCO2, Generation)) %>% 
                      dplyr::filter(Age %in% c(8, 14)) %>% 
                      tidyr::pivot_wider(names_from = Age, 
                                         values_from = proportion_rel_dstage,
                                         names_glue = "Age_{.name}") %>% # renmae new cols
                      dplyr::mutate(type       = "pre", # new col fo rpre
                                    delta.perc = Age_14 - Age_8) %>%  # sutract em
                      dplyr::mutate(delta.perc = case_when(delta.perc > 0 ~ 0, # positives make zero
                                                           .default = delta.perc),
                                    abs.delta.perc = abs(delta.perc)) %>% # now take the absolute value
                      na.omit() %>% 
                      dplyr::select(-c(Age_8,Age_14)) # no longer need these, remove to merge with post
  

Exp4_Post.calc <- Exp4_PrePost %>% 
                      dplyr::select(c(Age, proportion_rel_dstage, Tank, Larvae_pCO2,Generation)) %>% 
                      dplyr::filter(Age %in% c(14, 21)) %>% 
                      tidyr::pivot_wider(names_from = Age, 
                                         values_from = proportion_rel_dstage,
                                         names_glue = "Age_{.name}") %>% # renmae new cols
                      dplyr::mutate(type       = "post", # new col fo rpre
                                    delta.perc = Age_21 - Age_14) %>%  # sutract em
                      dplyr::mutate(delta.perc = case_when(delta.perc > 0 ~ 0, # positives make zero
                                                           .default = delta.perc),
                                    abs.delta.perc = abs(delta.perc))%>% # now take the absolute value
                      na.omit() %>% 
                      dplyr::select(-c(Age_14,Age_21)) # no longer need these, remove to merge with post


Exp4_PrePost.calc <- rbind(Exp4_Pre.calc,Exp4_Post.calc)
``` 


```{r Statistics for Pre and Post each Two-way anovas}
# load packages 
library(rcompanion)
library(FSA)

# merge datasets and assign gneration as a factor for the models
Exp2Exp4_PrePost.master            <- rbind(Exp2_PrePost.calc,Exp4_PrePost.calc)
Exp2Exp4_PrePost.master$Generation <- factor(Exp2Exp4_PrePost.master$Generation)


# delta pre - test the effects of larvae pCO2 and generation
TwoWay.Pre <- aov(lm(abs.delta.perc ~ Larvae_pCO2 * Generation, 
                     data = (Exp2Exp4_PrePost.master %>% dplyr::filter(type %in% 'pre'))
                  ))
shapiro.test(resid(TwoWay.Pre)) # 0.05353 - pass
leveneTest(TwoWay.Pre) # 0.7313 - pass
summary(TwoWay.Pre)

#                        Df Sum Sq Mean Sq F value   Pr(>F)    
# Larvae_pCO2             1 0.0023  0.0023   0.241    0.632    
# Generation              1 0.3412  0.3412  36.339 5.95e-05 ***
# Larvae_pCO2:Generation  1 0.0010  0.0010   0.107    0.749    
# Residuals              12 0.1127  0.0094 



# delta post - test the effects of larvae pCO2 and generation
TwoWay.Post <- aov(lm(abs.delta.perc ~ Larvae_pCO2 * Generation, 
                     data = (Exp2Exp4_PrePost.master %>% dplyr::filter(type %in% 'post'))
                  ))
shapiro.test(resid(TwoWay.Post)) # 0.007293 - non-normal
leveneTest(TwoWay.Post) # 0.6631 - pass
# * run non-parametric
SRH.Post <- scheirerRayHare(abs.delta.perc ~ Larvae_pCO2 * Generation, 
                     data = (Exp2Exp4_PrePost.master %>% dplyr::filter(type %in% 'post')))
#                        Df  Sum Sq      H p.value
# Larvae_pCO2             1 110.250 4.8640 0.02742
# Generation              1 166.667 7.3529 0.00670
# Larvae_pCO2:Generation  1   0.817 0.0360 0.84945
# Residuals              12  62.267
# posthoc dunn test for main effect of arvae pCO2

```

```{r Plots for Pre and Post}

Exp2Exp4_PrePost.summ <- Exp2Exp4_PrePost.master %>% 
                          dplyr::select(Generation,type,Larvae_pCO2,abs.delta.perc) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Generation,Larvae_pCO2,type) %>% 
                          dplyr::summarise(mean_delta.perc = mean(abs.delta.perc), 
                                           n           = n(),
                                           sd_delta.perc   = sd(abs.delta.perc),
                                           se_delta.perc   = sd_delta.perc/(sqrt(n)))

Exp2Exp4_PrePost.summ$type = factor(Exp2Exp4_PrePost.summ$type, levels =c("pre","post"))

DeltaPercent_barplot <- ggplot(Exp2Exp4_PrePost.summ) +
                          geom_errorbar(aes(x=Larvae_pCO2, 
                                             ymin=mean_delta.perc-se_delta.perc, 
                                             ymax=mean_delta.perc+se_delta.perc,
                                             group=Generation,
                                             fill=Larvae_pCO2), 
                                        position=position_dodge(.9),
                                        width=0, # removes the horizontal line
                                        colour="black",
                                        size=1) +
                          geom_bar(aes(x=Larvae_pCO2, 
                                       y=mean_delta.perc,
                                       fill=factor(Larvae_pCO2),
                                       group=Generation),
                                   position=position_dodge(.9),
                                   stat="identity",
                                   width = 0.75,
                                   alpha = 0.5) +
                          labs(title="Survival difference Pre/Post metamorphosis", 
                              x ="pCO2 Exposure", 
                              y = "Delta(abs(percent))") +
                           scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2"), 
                                               values=c("forestgreen","orange")) +
                          scale_x_discrete(labels=c("L", "M")) +
                          # scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
                          theme_classic() +
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                                axis.text=element_text(size=10),
                                legend.position="none") +
                          facet_wrap(~type)


capture.output(summary(TwoWay.Pre), file="Output/1_Survival_Growth/survival/Pre_Metamorphosis_TwoWayANOVA.doc")
capture.output(SRH.Post, file="Output/1_Survival_Growth/survival/Post_Metamorphosis_SRH.doc")

pdf("Output/1_Survival_Growth/survival/PrePost_Metamorphosis.pdf",width = 4, height = 4)
DeltaPercent_barplot
dev.off()
```














## Manuscript 4; Experiment 6

**output folder** = '4_F3_Full_factorial'

**note**: out ultimate objective with the survival data in this manuscript is to investigate puatative effects of parental envrionment and offspring
exposure on surivorship  - we reached metamorphosis at ?? DPF

### Experiment 4 F2s

```{r format F3 Experiment #6}
# create new columns for live and dead based on the count and proportion (of alive) to back calculate dead - these are integers! 
df_Exp6 <- df_Exp6 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(proportion = Percent_Survival/100,
                        Count_total = case_when(proportion == 0 ~ 0,
                                                proportion >0 ~
                                                 (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive)),# proportion is that of alive
                           Count_dead = Count_total - Count_alive)
```

```{r convert to binary F3 Experiment #6}

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
binary_Exp6_all     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_Exp6)) {
   count_alive <- round(df_Exp6[i,]$Count_alive/100) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp6[i,]$Count_dead/100)  # divide all counts by 1000, round to nearest
   
   #count_alive <- round(df_Exp6[i,]$Count_alive) # raw counts
   #count_dead  <- round(df_Exp6[i,]$Count_dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_Exp6[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_Exp6[i,]$Larvae_pCO2,
                            Generation  = df_Exp6[i,]$Generation,
                            Tank        = df_Exp6[i,]$Tank,  
                            Age         = df_Exp6[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp6_all <- rbind(binary_Exp6_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_Exp6_all$Tank) # "A" "B" "C" "D" "E" NA
binary_Exp6_all <- binary_Exp6_all %>% 
                    dplyr::mutate(Generation = 3,
                                  Tank_num = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3),
                                    pCO2_all = paste0(gsub('*.pCO2','',Parent_pCO2) , 
                                                      " x ", 
                                                      gsub('*.pCO2','',Larvae_pCO2)))  %>% 
                    dplyr::mutate(Tank = paste0(substr(Parent_pCO2,1,1),
                                                substr(Larvae_pCO2,1,1),
                                                '_',
                                                Tank_num)) %>% 
                    na.omit()
binary_Exp6_all$pCO2_all <- factor(binary_Exp6_all$pCO2_all, 
                                      levels =c("Low x Low", "Low x Moderate", "Low x High", 
                                                "Moderate x Low", "Moderate x Moderate", "Moderate x High", 
                                                "High x Low", "High x Moderate", "High x High"))









# (2) Mean counts alive and dead by tank - better representated for plots!
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_Exp6_MEANS <- df_Exp6 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Parent_pCO2, Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))

# call datafarme binary for plotting
binary_Exp6_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_Exp6_MEANS)) {
   count_alive <- round(df_Exp6_MEANS[i,]$mean_Count_alive/100) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp6_MEANS[i,]$mean_Count_dead/100)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 4, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2', 'Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_Exp6_MEANS[i,]$Parent_pCO2, 
                            Larvae_pCO2 = df_Exp6_MEANS[i,]$Larvae_pCO2,
                            Age         = df_Exp6_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),4] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),4] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp6_MEANS <- rbind(binary_Exp6_MEANS,loopDF) #bind to a cumulative list dataframe
  
}

binary_Exp6_MEANS <- binary_Exp6_MEANS %>% 
                      dplyr::mutate(Generation = 3,
                                    pCO2_all = paste0(gsub('*.pCO2','',Parent_pCO2) , 
                                                      " x ", 
                                                      gsub('*.pCO2','',Larvae_pCO2))) %>%
                    na.omit()


binary_Exp6_MEANS$pCO2_all <- factor(binary_Exp6_MEANS$pCO2_all, 
                                      levels =c("Low x Low", "Low x Moderate", "Low x High", 
                                                "Moderate x Low", "Moderate x Moderate", "Moderate x High", 
                                                "High x Low", "High x Moderate", "High x High"))


```

**note**: review the google doc here - https://docs.google.com/spreadsheets/d/1D_cT3HZDhmo5BltCnwcf2F97o5gy1HDDKXfYpYZwmbI/edit#gid=229896546
records indivcate that spat (post metamorposis) were counted on day 16, however only larvae on day 12, lets truncate data <=12 for larval survival 
```{r filter larval period F3 Experiment #6}
unique(binary_Exp6_larvae$Age) #  2  5  7 12 16
binary_Exp6_larvae      <- binary_Exp6_all %>% filter(Age <=12) # 

binary_Exp6_larvaeMEANS <- binary_Exp6_MEANS %>% filter(Age <=12) #
```

**note** this is a full factorial experiment parent x offspring pCO2, we can not use an interaction term in the surv object 
but we can use it additively corresponding with different kaplan meier. Before moving forward, lets also select by parent treatment 
to run separate surv objects andhazard analysis 
```{r filter by parental exposure for F3 Experiment #6}
binary_Exp6_larvae_LOW  <- binary_Exp6_larvae %>% dplyr::filter(Parent_pCO2 %in% 'Low pCO2')
binary_Exp6_larvae_MOD  <- binary_Exp6_larvae %>% dplyr::filter(Parent_pCO2 %in% 'Moderate pCO2')
binary_Exp6_larvae_HIGH <- binary_Exp6_larvae %>% dplyr::filter(Parent_pCO2 %in% 'High pCO2')


binary_Exp6_larvaeMEANS_LOW  <- binary_Exp6_larvaeMEANS %>% dplyr::filter(Parent_pCO2 %in% 'Low pCO2')
binary_Exp6_larvaeMEANS_MOD  <- binary_Exp6_larvaeMEANS %>% dplyr::filter(Parent_pCO2 %in% 'Moderate pCO2')
binary_Exp6_larvaeMEANS_HIGH <- binary_Exp6_larvaeMEANS %>% dplyr::filter(Parent_pCO2 %in% 'High pCO2')
```

### create surv object 

* 'all' meaning we kept the tank resolution to convert data to binary, assumes we can use 1|Tank in the surv model, else this is invalid
* 'means' meaning we had one vale based on mean counts live and dead for each treatment to convert to binary 

**note** from my experience with these data thus far, it appreas the tank random effect is not compatible, consolidating data to a mean
works well, confirm this
```{r surv objects F3 Experiment #6}
unique(binary_Exp6_larvaeMEANS$pCO2_all)
# all
survExp6_all   <- survfit2(Surv(Age, Count_dead) ~ pCO2_all, data = binary_Exp6_larvae)

# all based on parental 
survExp6_all.low   <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvae_LOW)
survExp6_all.mod   <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvae_MOD)
survExp6_all.high  <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvae_HIGH)

# means 
survExp6_means <- survfit2(Surv(Age, Count_dead) ~ pCO2_all, data = binary_Exp6_larvaeMEANS)

# means based on parental
survExp6_means.low  <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvaeMEANS_LOW)
survExp6_means.mod  <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvaeMEANS_MOD)
survExp6_means.high <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvaeMEANS_HIGH)

```

```{r plot Kaplan Meier F3 Experiment #6}

Exp6_larvaeKaplanMeier.low <- survExp6_means.low %>% 
                                ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                                scale_color_manual(values = c("forestgreen","orange","purple")) +
                                scale_fill_manual(values = c("forestgreen","orange","purple")) +
                                labs(
                                  x = "Days",
                                  y = "Overall survival probability"
                                ) + 
                                add_confidence_interval() +
                                add_risktable() +
                                add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                                scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                               # add_pvalue(location  = "annotation", x = 8.5)

Exp6_larvaeKaplanMeier.mod <- survExp6_means.mod %>% 
                                ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                                scale_color_manual(values = c("forestgreen","orange","purple")) +
                                scale_fill_manual(values = c("forestgreen","orange","purple")) +
                                labs(
                                  x = "Days",
                                  y = "Overall survival probability"
                                ) + 
                                add_confidence_interval() +
                                add_risktable() +
                                add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                                scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                               # add_pvalue(location  = "annotation", x = 8.5)

Exp6_larvaeKaplanMeier.high <- survExp6_means.high %>% 
                                ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                                scale_color_manual(values = c("forestgreen","orange","purple")) +
                                scale_fill_manual(values = c("forestgreen","orange","purple")) +
                                labs(
                                  x = "Days",
                                  y = "Overall survival probability"
                                ) + 
                                add_confidence_interval() +
                                add_risktable() +
                                add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                                scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                               # add_pvalue(location  = "annotation", x = 8.5)


# plot the surv data
pdf("Output/4_F3_Full_factorial/Expeirment6_Survival_KaplanMeier_LowParent.pdf", width=5, height = 4)
Exp6_larvaeKaplanMeier.low
dev.off()

pdf("Output/4_F3_Full_factorial/Expeirment6_Survival_KaplanMeier_ModParent.pdf", width=5, height = 4)
Exp6_larvaeKaplanMeier.mod
dev.off()

pdf("Output/4_F3_Full_factorial/Expeirment6_Survival_KaplanMeier_HighParent.pdf", width=5, height = 4)
Exp6_larvaeKaplanMeier.high
dev.off()


ggarrange(Exp6_larvaeKaplanMeier.low,Exp6_larvaeKaplanMeier.mod,Exp6_larvaeKaplanMeier.high,ncol=3)
```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).
* We get the log-rank p-value using the survdiff function
```{r log-rank test survdiff F4 Experiment #6}


Exp4_logrank_test.all <- survdiff(formula = Surv(Age, Count_dead) ~ pCO2_all, data = binary_Exp6_larvaeMEANS)
#                                  N Observed Expected (O-E)^2/E (O-E)^2/V
# pCO2_all=Low x Low           15826    14156    13918    4.0546     8.999
# pCO2_all=Low x Moderate      15850    14286    13931    9.0517    20.085
# pCO2_all=Low x High          15818    14311    13921   10.9158    24.237
# pCO2_all=Moderate x Low      15799    13741    13904    1.9024     4.223
# pCO2_all=Moderate x Moderate 15810    13739    13909    2.0815     4.620
# pCO2_all=Moderate x High     15760    13907    13882    0.0453     0.101
# pCO2_all=High x Low          15819    13737    13915    2.2757     5.051
# pCO2_all=High x Moderate     15788    13449    13896   14.3770    31.916
# pCO2_all=High x High         15782    13846    13896    0.1788     0.397


Exp4_logrank_test.low <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvaeMEANS_LOW)
#                               N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      15826    14156    14246    0.5664     1.686
# Larvae_pCO2=Moderate pCO2 15850    14286    14259    0.0523     0.156
# Larvae_pCO2=High pCO2     15818    14311    14248    0.2744     0.817

Exp4_logrank_test.mod <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvaeMEANS_MOD)
#                               N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      15799    13741    13801     0.261     0.776
# Larvae_pCO2=Moderate pCO2 15810    13739    13806     0.329     0.981
# Larvae_pCO2=High pCO2     15760    13907    13780     1.178     3.507

Exp4_logrank_test.high <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp6_larvaeMEANS_HIGH)
#                               N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      15819    13737    13690     0.163     0.477
# Larvae_pCO2=Moderate pCO2 15788    13449    13671     3.610    10.566
# Larvae_pCO2=High pCO2     15782    13846    13671     2.239     6.553
```

# The Cox regression model
* We may want to quantify an effect size for a single variable, or include **more than one variable** into a regression model to account for the effects of multiple variables.
* The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.

# Run several mdoels 

- mixed models - theseinclude all data with the Tank replicate (not the mean binary data!)
  
  * all - use 'pCO2_all' to run 
  
  * additive model - Parent_pCO2 and Larvae_pCO2 to explore main effects

### Mixed model - all
```{r cox & hazard ratio -  F3 Experiment #6 - pCO2 all mixed model}

binary_Exp6_larvae      <- droplevels(binary_Exp6_larvae) # clean empty levels to run coxme without error

# all - meaning that tank was kept, my gut is that the mean data is better suited as tank seems unaddressed here..
dependent_os        <- "Surv(Age, Count_dead)"
explanatory         <- c("pCO2_all", "frailty(Tank)")
explanatory_multi   <- c("Parent_pCO2","Larvae_pCO2", "frailty(Tank)")

binary_Exp6_larvae$Tank_num <- as.numeric(binary_Exp6_larvae$Tank)

Exp6_coxph_test         <- coxph(formula = Surv(Age, Count_dead) ~ # using coxpH and Tank as random factor using 'frailty'
                                   pCO2_all, # exclude +frailty(Tank) to be able to generate hazard plot
                                 data = binary_Exp6_larvae)
#                                  coef exp(coef)  se(coef)       z        p
# pCO2_allLow x Moderate       0.011008  1.011069  0.006847   1.608  0.10789
# pCO2_allLow x High           0.014876  1.014988  0.006844   2.174  0.02973
# pCO2_allModerate x Low      -0.035725  0.964905  0.006914  -5.167 2.38e-07
# pCO2_allModerate x Moderate -0.032332  0.968185  0.006914  -4.676 2.92e-06
# pCO2_allModerate x High     -0.007210  0.992816  0.006893  -1.046  0.29556
# pCO2_allHigh x Low          -0.044415  0.956557  0.006915  -6.423 1.34e-10
# pCO2_allHigh x Moderate     -0.074028  0.928645  0.006953 -10.647  < 2e-16
# pCO2_allHigh x High         -0.020382  0.979824  0.006901  -2.954  0.00314


# Fit a mixed effects Cox model

# # run with finalfit to verify HR to the coxpH run
binary_Exp6_larvae %>%  # same as above, using finalfit to output, includes frailty
    finalfit(dependent_os, explanatory, add_dependent_label = FALSE)

 #  label              levels          all          HR (univariable) HR (multivariable)
 #      pCO2_all           Low x Low 47483 (11.1)                         -                  -
 #                    Low x Moderate 47555 (11.1) 1.01 (1.00-1.02, p=0.108)   NA (NA-NA, p=NA)
 #                        Low x High 47455 (11.1) 1.01 (1.00-1.03, p=0.030)   NA (NA-NA, p=NA)
 #                    Moderate x Low 47398 (11.1) 0.96 (0.95-0.98, p<0.001)   NA (NA-NA, p=NA)
 #               Moderate x Moderate 47434 (11.1) 0.97 (0.96-0.98, p<0.001)                  -
 #                   Moderate x High 47278 (11.1) 0.99 (0.98-1.01, p=0.296)   NA (NA-NA, p=NA)
 #                        High x Low 47458 (11.1) 0.96 (0.94-0.97, p<0.001)   NA (NA-NA, p=NA)
 #                   High x Moderate 47361 (11.1) 0.93 (0.92-0.94, p<0.001)   NA (NA-NA, p=NA)
 #                       High x High 47348 (11.1) 0.98 (0.97-0.99, p=0.003)   NA (NA-NA, p=NA)
 # frailty(Tank)                                                          -                  -
 #          <NA>                <NA>         <NA>                      <NA>   NA (NA-NA, p=NA)


Exp6_coxph_test.mixed   <- coxme(Surv(Age, Count_dead) ~ # using coxme, notice ther results are the same as coxpH
                                   pCO2_all + (1 | Tank), # note we cannot make a hazrd ratio plot using coxme
                                 data = binary_Exp6_larvae)
# Model:  Surv(Age, Count_dead) ~ pCO2_all + (1 | Tank) 
# Fixed coefficients
#                                     coef exp(coef)   se(coef)     z       p
# pCO2_allLow x Moderate       0.010990968 1.0110516 0.01453067  0.76 4.5e-01
# pCO2_allLow x High           0.014817140 1.0149275 0.01452938  1.02 3.1e-01
# pCO2_allModerate x Low      -0.035791327 0.9648416 0.01456265 -2.46 1.4e-02
# pCO2_allModerate x Moderate -0.032436486 0.9680839 0.01456270 -2.23 2.6e-02
# pCO2_allModerate x High     -0.007349739 0.9926772 0.01455284 -0.51 6.1e-01
# pCO2_allHigh x Low          -0.044452742 0.9565208 0.01456290 -3.05 2.3e-03
# pCO2_allHigh x Moderate     -0.074084576 0.9285932 0.01458094 -5.08 3.8e-07
# pCO2_allHigh x High         -0.020481000 0.9797273 0.01455627 -1.41 1.6e-01
# 
# Random effects
#  Group Variable  Std Dev      Variance    
# Tank  Intercept 0.0156969373 0.0002463938


# Hazard ratio plots
Exp6_hazardratio_plotMIXED <- binary_Exp6_larvae %>% 
                                hr_plot(dependent_os, explanatory)

Exp6_hazardratio_plot      <- ggforest(Exp6_coxph_test, data = binary_Exp6_larvae)

ggarrange(Exp6_hazardratio_plotMIXED,Exp6_hazardratio_plot, ncol = 2)


# output yo stuff homie!
capture.output(Exp6_coxph_test.mixed,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_mixedmodel.doc")

pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_mixed_all.pdf", width = 6, height = 6) # three figures
Exp6_hazardratio_plot
dev.off()



```

### Mixed model - additive (main effects of parent and larvae pCO2)
```{r cox & hazard ratio -  F3 Experiment #6 - additive mixed model}
library(survminer) # ggforect hazard ratio plot
library(finalfit)


binary_Exp6_larvae$Parent_pCO2 <- as.factor(binary_Exp6_larvae$Parent_pCO2)
binary_Exp6_larvae$Larvae_pCO2 <- as.factor(binary_Exp6_larvae$Larvae_pCO2)


# all - meaning that tank was kept, my gut is that the mean data is better suited as tank seems unaddressed here..
dependent_os  <- "Surv(Age, Count_dead)"
explanatory   <- c("Parent_pCO2", "Larvae_pCO2", "frailty(Tank)")

Exp6_coxph_test         <- coxph(formula = Surv(Age, Count_dead) ~ # using coxpH and Tank as random factor using 'frailty'
                                   Parent_pCO2 + Larvae_pCO2, # exclude +frailty(Tank) to be able to generate hazard plot
                                 data = binary_Exp6_larvae)
#                               coef exp(coef)  se(coef)       z        p
# Parent_pCO2Moderate pCO2 -0.033674  0.966886  0.003981  -8.458  < 2e-16
# Parent_pCO2High pCO2     -0.054783  0.946690  0.003990 -13.729  < 2e-16
# Parent_pCO2None                 NA        NA  0.000000      NA       NA
# Larvae_pCO2Moderate pCO2 -0.004812  0.995199  0.004005  -1.201     0.23
# Larvae_pCO2High pCO2      0.022370  1.022622  0.003992   5.604 2.09e-08



# Fit a mixed effects Cox model

# # run with finalfit to verify HR to the coxpH run
binary_Exp6_larvae %>%  # same as above, using finalfit to output, includes frailty
    finalfit(dependent_os, explanatory, add_dependent_label = FALSE)
#     label        levels           all          HR (univariable) HR (multivariable)
# Parent_pCO2      Low pCO2 142493 (33.4)                         -                  -
#             Moderate pCO2 142110 (33.3) 0.97 (0.96-0.97, p<0.001)   NA (NA-NA, p=NA)
#                 High pCO2 142167 (33.3) 0.95 (0.94-0.95, p<0.001)   NA (NA-NA, p=NA)
# Larvae_pCO2      Low pCO2 142339 (33.4)                         -                  -
#             Moderate pCO2 142350 (33.4) 1.00 (0.99-1.00, p=0.222)   NA (NA-NA, p=NA)
#                 High pCO2 142081 (33.3) 1.02 (1.01-1.03, p<0.001)   NA (NA-NA, p=NA)

# run the coxme model, report this in the manuscript
binary_Exp6_larvae      <- droplevels(binary_Exp6_larvae) # clean empty levels to run coxme without error
Exp6_coxph_test.mixed   <- coxme(Surv(Age, Count_dead) ~ # using coxme, notice ther results are the same as coxpH
                                   Parent_pCO2 + Larvae_pCO2 + (1 | Tank), # note we cannot make a hazrd ratio plot using coxme
                                 data = binary_Exp6_larvae)
# Model:  Surv(Age, Count_dead) ~ Parent_pCO2 + Larvae_pCO2 + (1 | Tank) 
# Fixed coefficients
#                                  coef exp(coef)    se(coef)     z       p
# Parent_pCO2Moderate pCO2 -0.033807387 0.9667577 0.009492169 -3.56 3.7e-04
# Parent_pCO2High pCO2     -0.054920227 0.9465607 0.009495918 -5.78 7.3e-09
# Larvae_pCO2Moderate pCO2 -0.005038845 0.9949738 0.009502506 -0.53 6.0e-01
# Larvae_pCO2High pCO2      0.022392300 1.0226449 0.009496618  2.36 1.8e-02
# 
# Random effects
#  Group Variable  Std Dev      Variance    
#  Tank  Intercept 0.0182788805 0.0003341175


# Hazard ratio
Exp6_hazardratio_plotMIXED <- binary_Exp6_larvae %>% 
                                hr_plot(dependent_os, explanatory)

Exp6_hazardratio_plot      <- ggforest(Exp6_coxph_test, data = binary_Exp6_larvae) # report this one


# output yo stuff homie!
capture.output(Exp6_coxph_test.mixed,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_mixedmodel_maineffects.doc")

pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_mixed_maineffects.pdf", width = 6, height = 6) # three figures
Exp6_hazardratio_plot
dev.off()

```

```{r cox & hazard ratio -  F3 Experiment #6 - by parental LOW mixed model}

# for the all model (inluing Tank rnaodm) we need the fraility assignment of tank
# aletnatively can run come however cannot make a hazard ratio plot from the output of coxme
dependent_os  <- "Surv(Age, Count_dead)"
explanatory   <- c("Larvae_pCO2", "frailty(Tank)")


# Low Parents :::::::::::::::::::::::::::::::
# reference is low exposure for low parents
binary_Exp6_larvae_LOW$Larvae_pCO2 <- factor(binary_Exp6_larvae_LOW$Larvae_pCO2, levels =c('Low pCO2', 'Moderate pCO2', 'High pCO2'))

Exp6_coxph_test.low  <- coxph(formula = Surv(Age, Count_dead) ~ # not indcluding the random Tank effect
                                   Larvae_pCO2, 
                                 data = binary_Exp6_larvae_LOW)
#                              coef exp(coef) se(coef)     z      p
# Larvae_pCO2Moderate pCO2 0.011000  1.011061 0.006847 1.607 0.1081
# Larvae_pCO2High pCO2     0.014871  1.014982 0.006844 2.173 0.0298

binary_Exp6_larvae_LOW %>%  # same as above, using finalfit to output, includes frailty
    finalfit(dependent_os, explanatory, add_dependent_label = FALSE)
   #       label        levels          all          HR (univariable) HR (multivariable)
   # Larvae_pCO2      Low pCO2 47483 (33.3)                         -                  -
   #             Moderate pCO2 47555 (33.4) 1.01 (1.00-1.02, p=0.108)   NA (NA-NA, p=NA)
   #                 High pCO2 47455 (33.3) 1.01 (1.00-1.03, p=0.030)   NA (NA-NA, p=NA)

Exp6_coxph_test.lowmixed    <- coxme(Surv(Age, Count_dead) ~ Larvae_pCO2 + (1 | Tank), data = binary_Exp6_larvae_LOW)
# Model:  Surv(Age, Count_dead) ~ Larvae_pCO2 + (1 | Tank) 
# Fixed coefficients
#                                coef exp(coef)   se(coef)    z    p
# Larvae_pCO2Moderate pCO2 0.01099086  1.011051 0.01083440 1.01 0.31
# Larvae_pCO2High pCO2     0.01483468  1.014945 0.01083265 1.37 0.17
# 
# Random effects
#  Group Variable  Std Dev      Variance    
#  Tank  Intercept 0.0102839485 0.0001057596

Exp6_hazardratio_plot.low   <- ggforest(Exp6_coxph_test.low, data = binary_Exp6_larvae_LOW)

pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_mixed_LOW.pdf", width = 8, height = 6) # three figures
Exp6_hazardratio_plot.low
dev.off()

```

```{r cox & hazard ratio -  F3 Experiment #6 - by parental MOD mixed model}

# Moderate Parents :::::::::::::::::::::::::::::::
# reference is moderate exposure 
binary_Exp6_larvae_MOD$Larvae_pCO2 <- factor(binary_Exp6_larvae_MOD$Larvae_pCO2, 
                                             levels =c('Moderate pCO2', 'Low pCO2', 'High pCO2'))


Exp6_coxph_test.mod  <- coxph(formula = Surv(Age, Count_dead) ~ # not indcluding the random Tank effect
                                   Larvae_pCO2, 
                                 data = binary_Exp6_larvae_MOD)
#                               coef exp(coef)  se(coef)      z     p
# Larvae_pCO2Moderate pCO2 -0.008051  0.991981  0.020634 -0.390 0.696
# Larvae_pCO2High pCO2      0.031963  1.032479  0.020649  1.548 0.122

binary_Exp6_larvae_MOD %>%  # same as above, using finalfit to output, includes frailty
    finalfit(dependent_os, explanatory, add_dependent_label = FALSE)
   #       label        levels         all          HR (univariable)        HR (multivariable)
   # Larvae_pCO2      Low pCO2 5283 (33.2)                         -                         -
   #             Moderate pCO2 5352 (33.6) 0.99 (0.95-1.03, p=0.696) 0.99 (0.95-1.03, p=0.666)
   #                 High pCO2 5274 (33.2) 1.03 (0.99-1.08, p=0.122) 1.03 (0.99-1.07, p=0.130)

Exp6_coxph_test.modmixed    <- coxme(Surv(Age, Count_dead) ~ Larvae_pCO2 + (1 | Tank), data = binary_Exp6_larvae_MOD)
#                              coef exp(coef)   se(coef)     z    p
# Larvae_pCO2Low pCO2  -0.003758619 0.9962484 0.02106181 -0.18 0.86
# Larvae_pCO2High pCO2  0.025496459 1.0258243 0.02105496  1.21 0.23

Exp6_hazardratio_plot.mod   <- ggforest(Exp6_coxph_test.mod, data = binary_Exp6_larvae_MOD)

pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_mixed_MOD.pdf", width = 8, height = 6) # three figures
Exp6_coxph_test.mod
dev.off()


```

```{r cox & hazard ratio -  F3 Experiment #6 - by parental HIGH mixed model}


# High Parents :::::::::::::::::::::::::::::::
# referencig high exposure for highi parents 
binary_Exp6_larvae_HIGH$Larvae_pCO2 <- factor(binary_Exp6_larvae_HIGH$Larvae_pCO2, 
                                              levels =c('High pCO2', 'Low pCO2', 'Moderate pCO2'))


Exp6_coxph_test.high  <- coxph(formula = Surv(Age, Count_dead) ~ # not indcluding the random Tank effect
                                   Larvae_pCO2, 
                                 data = binary_Exp6_larvae_HIGH)
#                               coef exp(coef)  se(coef)      z        p
# Larvae_pCO2Low pCO2      -0.023295  0.976974  0.006953 -3.350 0.000808
# Larvae_pCO2Moderate pCO2 -0.052559  0.948798  0.006992 -7.517  5.6e-14

binary_Exp6_larvae_HIGH %>%  # same as above, using finalfit to output, includes frailty
    finalfit(dependent_os, explanatory, add_dependent_label = FALSE)
   #       label        levels          all          HR (univariable) HR (multivariable)
   # Larvae_pCO2     High pCO2 47348 (33.3)                         -                  -
   #                  Low pCO2 47458 (33.4) 0.98 (0.96-0.99, p=0.001)   NA (NA-NA, p=NA)
   #             Moderate pCO2 47361 (33.3) 0.95 (0.94-0.96, p<0.001)   NA (NA-NA, p=NA)

Exp6_coxph_test.highmixed    <- coxme(Surv(Age, Count_dead) ~ Larvae_pCO2 + (1 | Tank), data = binary_Exp6_larvae_HIGH)
#                                 coef exp(coef)   se(coef)     z       p
# Larvae_pCO2Low pCO2      -0.02325405 0.9770142 0.01167296 -1.99 4.6e-02
# Larvae_pCO2Moderate pCO2 -0.05252814 0.9488276 0.01169589 -4.49 7.1e-06

Exp6_hazardratio_plot.high   <- ggforest(Exp6_coxph_test.high, data = binary_Exp6_larvae_HIGH)


pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_mixed_HIGH.pdf", width = 8, height = 6) # three figures
Exp6_hazardratio_plot.high
dev.off()



# output yo stuff homie!
capture.output(Exp6_coxph_test.mixed,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_mixedmodel.doc")
capture.output(Exp6_coxph_test.lowmixed,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_low.doc")
capture.output(Exp6_coxph_test.modmixed,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_mod.doc")
capture.output(Exp6_coxph_test.highmixed,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_high.doc")


pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_allplots.pdf", width = 15, height = 10) # three figures
ggarrange(Exp6_hazardratio_plot,
          (ggarrange(Exp6_hazardratio_plot.low, 
                     Exp6_hazardratio_plot.mod,
                     Exp6_hazardratio_plot.high,nrow=3)),
          ncol=2
)
dev.off()

















# means - one value per treatment, (no tank included) 
Exp6_coxph_testMEANS         <- coxph(formula = Surv(Age, Count_dead) ~ 
                                        pCO2_all, 
                                      data = binary_Exp6_larvaeMEANS)
#                                   coef  exp(coef)   se(coef)      z      p
# pCO2_allHigh x Low          -0.0008721  0.9991283  0.0353398 -0.025 0.9803
# pCO2_allHigh x Moderate     -0.0267778  0.9735776  0.0353185 -0.758 0.4483
# pCO2_allLow x High           0.0439804  1.0449618  0.0356012  1.235 0.2167
# pCO2_allLow x Low            0.0271771  1.0275498  0.0353097  0.770 0.4415
# pCO2_allLow x Moderate       0.0511495  1.0524802  0.0352876  1.450 0.1472
# pCO2_allModerate x High      0.0592975  1.0610909  0.0353574  1.677 0.0935
# pCO2_allModerate x Low       0.0258839  1.0262218  0.0354605  0.730 0.4654
# pCO2_allModerate x Moderate  0.0172404  1.0173899  0.0353331  0.488 0.6256
Exp6_hazardratio_plotMEANS   <- ggforest(Exp6_coxph_testMEANS, data = binary_Exp6_larvaeMEANS)
Exp6_coxph_test_tableMEANS   <- coxph(Surv(Age, Count_dead) ~ pCO2_all, 
                                      data = binary_Exp6_larvaeMEANS) %>%  tbl_regression(exp = TRUE) 


# means based on parental
Exp6_coxph_testMEANS.low         <- coxph(formula = Surv(Age, Count_dead) ~ 
                                   Larvae_pCO2, 
                                 data = binary_Exp6_larvaeMEANS_LOW)
Exp6_hazardratio_plotMEANS.low   <- ggforest(Exp6_coxph_testMEANS.low, data = binary_Exp6_larvaeMEANS_LOW)
Exp6_coxph_test_tableMEANS.low   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                 data = binary_Exp6_larvaeMEANS_LOW) %>%  tbl_regression(exp = TRUE) 

Exp6_coxph_testMEANS.mod         <- coxph(formula = Surv(Age, Count_dead) ~ 
                                   Larvae_pCO2, 
                                 data = binary_Exp6_larvaeMEANS_MOD)
Exp6_hazardratio_plotMEANS.mod   <- ggforest(Exp6_coxph_testMEANS.mod, data = binary_Exp6_larvaeMEANS_MOD)
Exp6_coxph_test_tableMEANS.mod   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                 data = binary_Exp6_larvaeMEANS_MOD) %>%  tbl_regression(exp = TRUE) 

Exp6_coxph_testMEANS.high         <- coxph(formula = Surv(Age, Count_dead) ~ 
                                   Larvae_pCO2, 
                                 data = binary_Exp6_larvaeMEANS_HIGH)
Exp6_hazardratio_plotMEANS.high   <- ggforest(Exp6_coxph_testMEANS.high, data = binary_Exp6_larvaeMEANS_HIGH)
Exp6_coxph_test_tableMEANS.high   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                 data = binary_Exp6_larvaeMEANS_HIGH) %>%  tbl_regression(exp = TRUE) 



# output yo stuff homie!
capture.output(Exp6_coxph_testMEANS,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_means_all.doc")
capture.output(Exp6_coxph_testMEANS.low,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_means_low.doc")
capture.output(Exp6_coxph_testMEANS.mod,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_means_mod.doc")
capture.output(Exp6_coxph_testMEANS.high,file="Output/4_F3_Full_factorial/survival/Experiment6_Survival_CoxReg_means_high.doc")


pdf("Output/4_F3_Full_factorial/survival/Experiment6_HazardRatio_means.pdf", width = 12, height = 9) # three figures
ggarrange(Exp6_hazardratio_plotMEANS,
          (ggarrange(Exp6_hazardratio_plotMEANS.low, 
                     Exp6_hazardratio_plotMEANS.mod,
                     Exp6_hazardratio_plotMEANS.high,nrow=3)),
          ncol=2
)
dev.off()

?coxme
```





























## Estimating x-time survival
* per 8 days from day 2 to 18 post fertilization
```{r}

F1_surv_prob <- s1_all %>% 
                          tbl_survfit(
                            times = 18,
                            label_header = "**survival to 20 days post-fertilization (95% CI)**")
F1_surv_prob$table_body$stat_1[2]
F1_surv_prob$table_body$stat_1[3]


# for loop every 50 days between 100 - 450 days post fertilization to a summary table!
x <- seq.int(2,18, by = 8) # call the integer
summarytable <- data.frame() # summary table 
for (i in 1:length(x)) { # for loop
  survtable <- s1_all %>% tbl_survfit(times = x[i])
  survtable$table_body$stat_1[2]
  survtable$table_body$stat_1[3]
  loopDF  <- data.frame(matrix(0,ncol = 4, nrow = 1)) 
  colnames(loopDF) = c('Age','LowpCO2_Probabilty','ModpCO2_Probabilty','Diff')
  loopDF  <- loopDF %>% 
        dplyr::mutate(
                      Age                = x[i],
                      LowpCO2_Probabilty = survtable$table_body$stat_1[2], 
                      ModpCO2_Probabilty = survtable$table_body$stat_1[3],
                      Diff               = abs(
                                           as.numeric(substr(survtable$table_body$stat_1[2],1,2)) -
                                           as.numeric(substr(survtable$table_body$stat_1[3],1,2)) 
                                           )
                      ) 
  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  summarytable <- rbind(summarytable,loopDF) #bind to a cumulative list dataframe
}
```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).






```{r}
# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_all  <- Surv(time = as.numeric(binary_Exp2_all$Age), 
                     event = as.numeric(binary_Exp2_all$Count_dead), type = "right")

# plot the surv data
ggsurvplot(survfit(surv_F1_all ~ Larvae_pCO2, binary_Exp2_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_means_all  <- Surv(time = as.numeric(binary_Exp2_MEANS$Age), 
                     event = as.numeric(binary_Exp2_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_Exp2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

s1 <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
str(s1)

View(binary_df_F1_MEANS)
s1 %>% 
  ggsurvfit(linetype_aes = TRUE) +
  scale_color_manual(values = c("forestgreen","orange","purple")) +
  scale_fill_manual(values = c("forestgreen","orange","purple")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )

?ggsurvfit
s1 %>% 
  ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
  scale_color_manual(values = c("forestgreen","orange","purple")) +
  scale_fill_manual(values = c("forestgreen","orange","purple")) +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable() +
  add_risktable_strata_symbol(symbol = "\U25CF", size = 10) +
  scale_ggsurvfit(x_scales = list(breaks = 0:10)) +
  add_pvalue(location  = "annotation", x = 8.5)

# use gtsummary 
s1 %>% 
  tbl_survfit(
    times = 10,
    label_header = "**survival to 10 days post-fertilization (95% CI)**"
  )
s1 %>% 
  tbl_survfit(
    probs = 0.5,
    label_header =  "**Median survival to 10 days post-fertilization (95% CI)**"
  )
s1
# Call: survfit(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
# 
#    2 observations deleted due to missingness 
#                              n events median 0.95LCL 0.95UCL
# Larvae_pCO2=Low pCO2      1844   1631      8       8       8
# Larvae_pCO2=Moderate pCO2 1460   1304      8       8       8
# Larvae_pCO2=High pCO2      805    762      4       4       4
survdiff(Surv(Age, Count_dead) ~ Larvae_pCO2 + (1|Tank), data = binary_df_F1_all)
binary_df_F1_all$Tank <- as.factor(as.numeric(binary_df_F1_all$Tank))
# Call:
# survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
# 
# n=4109, 2 observations deleted due to missingness.
# 
#                              N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      1844     1631     1917     42.67    144.62
# Larvae_pCO2=Moderate pCO2 1460     1304     1372      3.38      8.63
# Larvae_pCO2=High pCO2      805      762      408    307.29    500.45


coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS) %>% 
  tbl_regression(exp = TRUE) 


?survfit2
Surv(lung$time, lung$status)[1:10]
Surv(binary_df_F1_MEANS$Age, binary_df_F1_MEANS$Count_dead)[1:10]
head(lung[, c("time", "status", "sex")])

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F1_survival.pdf")
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_df_F1_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```
### Survival status 
* run the stats with the 'all' data and address the Tank as random effect
```{r F1 run coxme on surv stats}

head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)
#                               coef exp(coef)   se(coef)     z p
# Larvae_pCO2Moderate pCO2 0.1969625  1.217698 0.01697219 11.61 0
# Larvae_pCO2High pCO2     1.1161450  3.053062 0.02143035 52.08 0



#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```

```{r F1 run coxph on surv stats}
?coxme
head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)

#                               coef exp(coef)   se(coef)     z p
# Larvae_pCO2Moderate pCO2 0.1969625  1.217698 0.01697219 11.61 0
# Larvae_pCO2High pCO2     1.1161450  3.053062 0.02143035 52.08 0



#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
?glht
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```



```{r F2 build surv object and plot em}


# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_all  <- Surv(time = as.numeric(binary_df_F2_all$Age), 
                     event = as.numeric(binary_df_F2_all$Count_dead), type = "right")
# plot the surv data
ggsurvplot(survfit(surv_F2_all ~ Larvae_pCO2, binary_df_F2_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_means_all  <- Surv(time = as.numeric(binary_df_F2_MEANS$Age), 
                     event = as.numeric(binary_df_F2_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F2_survival.pdf")
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```

```{r F2 run coxme on surv stats}

head(binary_df_F2_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# omit the level high pCO2 because it is not actually present - to be a carry over from the full dataset?
is.na(binary_df_F2_all$Larvae_pCO2) <- binary_df_F2_all$Larvae_pCO2 == "High pCO2"
binary_df_F2_all$Larvae_pCO2 <- factor(binary_df_F2_all$Larvae_pCO2)
str(binary_df_F2_all) # Larvae_pCO2 now has two factors - correct!!

# now run the model
F2_survFIT_cox <- coxme(surv_F2_all ~ Larvae_pCO2 + (1|Tank), binary_df_F2_all)
summary(F2_survFIT_cox)

#                               coef exp(coef)   se(coef)   z      p
# Larvae_pCO2Moderate pCO2 0.1269179  1.135324 0.04536691 2.8 0.0051


#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)

glht(F2_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1269


```



# F1 and F2 responses across generations 
  * truncate data to day 14 that overlaps between both datasets 
  * plot and run stats 
```{r}


# filter and merge up to d14 (common between datasets)

# all data
unique(binary_df_F1_all$Age)
binary_df_F1_d10 <- binary_df_F1_all %>% dplyr::filter(!Age %in% c(14,18)) %>% na.omit()
unique(binary_df_F1_d10$Age) #  2  4  8 10
unique(binary_df_F2_all$Age)
binary_df_F2_d10 <- binary_df_F2_all %>% dplyr::filter(!Age %in% c(14,21)) %>% na.omit()
unique(binary_df_F2_d10$Age) # 2  6  8 10

binary_F1_F2_master   <- rbind(binary_df_F1_d10, binary_df_F2_d10) %>% dplyr::filter(!(Generation ==1 & 
                                                                                       Larvae_pCO2 %in% 'High pCO2'))
unique(binary_F1_F2_master$Larvae_pCO2)
binary_F1_F2_moderate <- binary_F1_F2_master %>% dplyr::filter(Larvae_pCO2 %in% 'Moderate pCO2')
binary_F1_F2_low      <- binary_F1_F2_master %>% dplyr::filter(Larvae_pCO2 %in% 'Low pCO2')

# by means
unique(binary_df_F2_MEANS$Age)
binary_df_F1_d10 <- binary_df_F1_MEANS %>% dplyr::filter(!Age %in% c(14,18)) %>% na.omit()
unique(binary_df_F1_d10$Age) #  2  4  8 10
unique(binary_df_F2_MEANS$Age)
binary_df_F2_d10 <- binary_df_F2_MEANS %>% dplyr::filter(!Age %in% c(14,21)) %>% na.omit()
unique(binary_df_F2_d10$Age) # 2  6  8 10

binary_F1_F2_masterMEANS   <- rbind(binary_df_F1_d10, binary_df_F2_d10) %>% dplyr::filter(!(Generation ==1 & 
                                                                                       Larvae_pCO2 %in% 'High pCO2'))
binary_F1_F2_moderateMEANS <- binary_F1_F2_masterMEANS %>% dplyr::filter(Larvae_pCO2 %in% 'Moderate pCO2')
binary_F1_F2_lowMEANS      <- binary_F1_F2_masterMEANS %>% dplyr::filter(Larvae_pCO2 %in% 'Low pCO2')



# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)


surv_F1_F2_allMEANS  <- Surv(time = as.numeric(binary_F1_F2_masterMEANS$Age), 
                     event = as.numeric(binary_F1_F2_masterMEANS$Count_dead), type = "right")


surv_F1_F2_all  <- Surv(time = as.numeric(binary_F1_F2_master$Age), 
                     event = as.numeric(binary_F1_F2_master$Count_dead), type = "right")
surv_F1_F2_modMEANS  <- Surv(time = as.numeric(binary_F1_F2_moderateMEANS$Age), 
                     event = as.numeric(binary_F1_F2_moderateMEANS$Count_dead), type = "right")
surv_F1_F2_lowMEANS  <- Surv(time = as.numeric(binary_F1_F2_lowMEANS$Age), 
                     event = as.numeric(binary_F1_F2_lowMEANS$Count_dead), type = "right")
# plot the surv data
# all 
ggsurvplot(survfit(surv_F1_F2_all ~ Generation+Larvae_pCO2, binary_F1_F2_masterMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("yellowgreen","orange","darkgreen","darkorange"), 
                             c(1,0.8,1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "Generation")
# low
# moderate
ggsurvplot(survfit(surv_F1_F2_modMEANS ~ Generation, binary_F1_F2_moderateMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("darkorange","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,
           xlab = "Age (days post fertilization)", 
           legend.title = "Generation")
# low
ggsurvplot(survfit(surv_F1_F2_lowMEANS ~ Generation, binary_F1_F2_lowMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("darkgreen","green"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "Generation")


# Estimate median survival time 
survfit(Surv(as.numeric(Generation), Larvae_pCO2) ~ 1, data = binary_F1_F2_master)
survfit(Surv(as.numeric(Generation), Larvae_pCO2) ~ 1, data = binary_F1_F2_master) %>% 
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median survival (95% CI)**"
  )

# statistics moderate
binary_F1_F2_master$Generation <- factor(binary_F1_F2_master$Generation)
str(binary_F1_F2_master) # Larvae_pCO2 now has two factors - correct!!
F1F2_survFIT_all_cox <- coxme(surv_F1_F2_all ~ Generation + (1|Tank), binary_F1_F2_master)
summary(F1F2_survFIT_Moderate_cox)


# as factor for the Gneration (the ID for this model)
binary_F1_F2_moderate$Generation <- factor(binary_F1_F2_moderate$Generation)
str(binary_F1_F2_moderate) # Larvae_pCO2 now has two factors - correct!!
# now run the model

survdiff(Surv(time, status) ~ sex, data = lung)


F1F2_survFIT_Moderate_cox <- coxme(surv_F1_F2_mod ~ Generation + (1|Tank), binary_F1_F2_moderate)
summary(F1F2_survFIT_Moderate_cox)
# Fixed coefficients
#                 coef exp(coef)   se(coef)     z       p
# Generation -0.208019 0.8121916 0.03484548 -5.97 2.4e-09
library(multcomp)
summary(glht(F1F2_survFIT_Moderate_cox, linfct = mcp(Generation = "Dunnett"),alternative = "greater"))
# Linear Hypotheses:
#            Estimate Std. Error z value Pr(>z)
# 2 - 1 <= 0 -0.20802    0.03485   -5.97      1



# statistics loq

# as factor for the Gneration (the ID for this model)
binary_F1_F2_low$Generation <- factor(binary_F1_F2_low$Generation)
str(binary_F1_F2_low) # Larvae_pCO2 now has two factors - correct!!
# now run the model
F1F2_survFIT_Low_cox <- coxme(surv_F1_F2_low ~ Generation + (1|Tank), binary_F1_F2_low)
summary(F1F2_survFIT_Low_cox)
# Fixed coefficients
#                   coef exp(coef)   se(coef)     z       p
# Generation2 -0.2371305 0.7888884 0.03939462 -6.02 1.8e-09
library(multcomp)
summary(glht(F1F2_survFIT_Low_cox, linfct = mcp(Generation = "Dunnett"),alternative = "greater"))
# Linear Hypotheses:
#            Estimate Std. Error z value Pr(>z)
# 2 - 1 <= 0 -0.23713    0.03939  -6.019      1


```
# F3 Larval Survival
## Plot it!
```{r F3 summary and plot}

stL_F3 <- summarySE(df_F3, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))
stL_F3 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Gen3_PlotMeanSE <- ggplot(stL_F3, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ 
                        geom_line()+
                        geom_point()+ 
                        geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                      position=position_dodge(.1)) +
                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                        theme(legend.position="right", legend.direction="vertical", 
                              legend.title = element_blank())+ theme_bw() + 
                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                        ggtitle("F3 larvae survival (Mean +- SE)") + 
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              panel.background = element_blank(), 
                              axis.line = element_line(colour = "black"))+
                        facet_wrap("Parent_pCO2")
Gen3_PlotMeanSE


# Summarise Percent Deformities for plotting 
Gen3_mean_survival <- df_F3 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Age, Percent_Survival) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Age, Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_surv = mean(Percent_Survival), 
                                           n           = n(),
                                           sd_Perc_surv   = sd(Percent_Survival),
                                           se_Perc_surv   = sd_Perc_surv/(sqrt(n)))

# Barplots 
unique(Gen3_mean_survival$Age)
Gen3_barplots_Age2 <- Gen3_mean_survival %>% dplyr::filter(Age ==2) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 2 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age5 <- Gen3_mean_survival %>% dplyr::filter(Age ==5) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 5 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age7 <- Gen3_mean_survival %>% dplyr::filter(Age ==7) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 7 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age12 <- Gen3_mean_survival %>% dplyr::filter(Age ==12) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 12 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age16 <- Gen3_mean_survival %>% dplyr::filter(Age ==16) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 16 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

# print the plot

pdf("Output/Survival/F3_survival_meanSE.pdf", width=10, height=5)
print(Gen3_PlotMeanSE_All)
dev.off()
# 
# 
# pdf("Output/Survival/PercSurvival_Gen3_2DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age2)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_5DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age5)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_7DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age7)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_12DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age12)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_16DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age16)
# dev.off()

```

```{r F3 build binary live and dead column}
#make a surv object to pass into coxme
df_F3 <- df_F3 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = 
                          (1/as.numeric(proportion))*as.numeric(Count_alive)) # proportion is that of alive


# Run two for loops 

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F3_all     <- data.frame() # start dataframe 
binary_df_F3_high    <- data.frame() # start dataframe 
binary_df_F3_mod     <- data.frame() # start dataframe 
binary_df_F3_low     <- data.frame() # start dataframe 
# run it 
for (i in 1:nrow(df_F3)) {
   #count_alive <- round(df_F3[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3[i,]$Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3[i,]$Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3[i,]$Larvae_pCO2,
                            Generation  = df_F3[i,]$Generation,
                            Tank        = df_F3[i,]$Tank,  
                            Age         = df_F3[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_all <- rbind(binary_df_F3_all,loopDF) #bind to a cumulative list dataframe
  # print(binary_df_F3) # print to monitor progress

     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_high <- rbind(binary_df_F3_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_mod <- rbind(binary_df_F3_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_low <- rbind(binary_df_F3_low,loopDF) #bind to a cumulative list dataframe
    }
      
  
}

binary_df_F3_all$All_pCO2 <- as.factor(paste0(binary_df_F3_all$Parent_pCO2, ' x ', binary_df_F3_all$Larvae_pCO2))


# (2) Mean counts alive and dead by tank - better presentative for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F3_MEANS <- df_F3 %>% 
  dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
  dplyr::group_by(Parent_pCO2, Larvae_pCO2, Age) %>% 
  dplyr::summarise(mean_Count_alive = mean(Count_alive), 
            mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F3_MEANS       <- data.frame() # start dataframe 
binary_df_F3_MEANS_high  <- data.frame() # start dataframe 
binary_df_F3_MEANS_mod   <- data.frame() # start dataframe 
binary_df_F3_MEANS_low   <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F3_MEANS)) {
   #count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 4, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3_MEANS[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F3_MEANS[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),4] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),4] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_MEANS <- rbind(binary_df_F3_MEANS,loopDF) #bind to a cumulative list dataframe
     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_MEANS_high <- rbind(binary_df_F3_MEANS_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_MEANS_mod <- rbind(binary_df_F3_MEANS_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_MEANS_low <- rbind(binary_df_F3_MEANS_low,loopDF) #bind to a cumulative list dataframe
  }
    
}
```

```{r F3 build surv object}
nrow(binary_df_F3_all)
# create surv object
# with tank included in the data frame (first for loop above)
surv_F3_all  <- Surv(time = as.numeric(binary_df_F3_all$Age), 
                     event = as.numeric(binary_df_F3_all$Count_dead), type = "right")

surv_F3_high <- Surv(time = as.numeric(binary_df_F3_high$Age), 
                     event = as.numeric(binary_df_F3_high$Count_dead), type = "right")

surv_F3_mod  <- Surv(time = as.numeric(binary_df_F3_mod$Age), 
                     event = as.numeric(binary_df_F3_mod$Count_dead), type = "right")

surv_F3_low  <- Surv(time = as.numeric(binary_df_F3_low$Age), 
                     event = as.numeric(binary_df_F3_low$Count_dead), type = "right")


F3_low_survplot <- ggsurvplot(survfit(surv_F3_low ~ Larvae_pCO2, binary_df_F3_low), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_mod_survplot <- ggsurvplot(survfit(surv_F3_mod ~ Larvae_pCO2, binary_df_F3_mod), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_high_survplot <- ggsurvplot(survfit(surv_F3_high ~ Larvae_pCO2, binary_df_F3_high), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_full_plot <- ggarrange(F3_low_survplot, F3_mod_survplot, F3_high_survplot, ncol = 3)
print(F3_full_plot)


# create surv object
# without tank included - means wihtin treatment SEPCIFICALLY FOR PLOTTING  second for loop above 
surv_F3_MEANS  <- Surv(time = as.numeric(binary_df_F3_MEANS$Age), 
                     event = as.numeric(binary_df_F3_MEANS$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_low$Parent_pCO2) <- binary_df_F3_MEANS_low$Parent_pCO2 == "None"
binary_df_F3_MEANS_low$Parent_pCO2 <- factor(binary_df_F3_MEANS_low$Parent_pCO2)
str(binary_df_F3_MEANS_low) # youll see now its fixed

surv_F3_MEANS_low  <- Surv(time = as.numeric(binary_df_F3_MEANS_low$Age), 
                     event = as.numeric(binary_df_F3_MEANS_low$Count_dead), type = "right")

is.na(binary_df_F3_MEANS_mod$Parent_pCO2) <- binary_df_F3_MEANS_mod$Parent_pCO2 == "None"
binary_df_F3_MEANS_mod$Parent_pCO2 <- factor(binary_df_F3_MEANS_mod$Parent_pCO2)
str(binary_df_F3_MEANS_mod) # youll see now its fixed

surv_F3_MEANS_mod  <- Surv(time = as.numeric(binary_df_F3_MEANS_mod$Age), 
                     event = as.numeric(binary_df_F3_MEANS_mod$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_high$Parent_pCO2) <- binary_df_F3_MEANS_high$Parent_pCO2 == "None"
binary_df_F3_MEANS_high$Parent_pCO2 <- factor(binary_df_F3_MEANS_high$Parent_pCO2)
str(binary_df_F3_MEANS_high) # youll see now its fixed

surv_F3_MEANS_high  <- Surv(time = as.numeric(binary_df_F3_MEANS_high$Age), 
                     event = as.numeric(binary_df_F3_MEANS_high$Count_dead), type = "right")

```

```{r F3 plot em}

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_all ~ Larvae_pCO2, binary_df_F3_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_MEANS ~ Larvae_pCO2, binary_df_F3_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_surival_low_history.pdf")
 ggsurvplot(survfit(surv_F3_MEANS_low ~ Larvae_pCO2, binary_df_F3_MEANS_low), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_moderate_history.pdf")
ggsurvplot(survfit(surv_F3_MEANS_mod ~ Larvae_pCO2, binary_df_F3_MEANS_mod), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_high_history.pdf")
# ggsurvplot(survfit(surv_F3_MEANS_high ~ Larvae_pCO2, binary_df_F3_MEANS_high), 
#            risk.table = F, pval = F , conf.int = TRUE,
#            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
#            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
#            legend.title = "treatment")
# dev.off()

```
run cox proportional test
include tank as a random effect
```{r}

# remove "none" and the level - for some reason this lingered, dont know why
str(binary_df_F3_all) # before running, youll see PArent_pCO2 has 4 levels - we only want 3!
is.na(binary_df_F3_all$Parent_pCO2) <- binary_df_F3_all$Parent_pCO2 == "None"
binary_df_F3_all$Parent_pCO2 <- factor(binary_df_F3_all$Parent_pCO2)
str(binary_df_F3_all) # youll see now its fixed

# now run the model

survFitME_all_tank <- coxme(surv_F3_all ~ Larvae_pCO2*Parent_pCO2 + (1|Tank), binary_df_F3_all)
summary(survFitME_all_tank) 
#                                                           coef exp(coef)   se(coef)     z
# Larvae_pCO2Moderate pCO2                          -0.025569607 0.9747545 0.01848915 -1.38
# Larvae_pCO2High pCO2                               0.009866940 1.0099158 0.01873860  0.53
# Parent_pCO2Moderate pCO2                          -0.037194870 0.9634884 0.01855713 -2.00
# Parent_pCO2High pCO2                              -0.076000227 0.9268160 0.01845043 -4.12
# Larvae_pCO2Moderate pCO2:Parent_pCO2Moderate pCO2  0.030348803 1.0308140 0.02614267  1.16
# Larvae_pCO2High pCO2:Parent_pCO2Moderate pCO2      0.008771872 1.0088105 0.02630283  0.33
# Larvae_pCO2Moderate pCO2:Parent_pCO2High pCO2      0.035694517 1.0363392 0.02605141  1.37
# Larvae_pCO2High pCO2:Parent_pCO2High pCO2          0.006232918 1.0062524 0.02611643  0.24
# p
#  1.7e-01
#      6.0e-01
#   4.5e-02
#     3.8e-05
#  2.5e-01
#    7.4e-01
#  1.7e-01
#   8.1e-01

#The standard deviation of the tank random effect is very small (0.015) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (527597), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 

#pairwise comparisons

library(multcomp)

glht(survFitME_all_tank, linfct = mcp(Parent_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0 -0.03719
# High pCO2 - Low pCO2 <= 0     -0.07600


```




# Stats
Lenght data do not meet the assumptions of noramlity even with an asin(sqrt()) transformation
```{r}
leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F1)
shapiro.test(df_F1$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F2)
shapiro.test(df_F2$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F3)
shapiro.test(df_F3$Percent_Survival_trans)
```

#Non parametric Stats
### F1 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F1)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F1,
              method="bh")

```

### F2 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F2)

res.kruskal <- kruskal.test(Percent_Survival ~ Larvae_pCO2, data = df_F2)
res.kruskal

```

### F3 Survival
```{r}


scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Parent_pCO2,
                data = df_F3)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F3,
              method="bh")

dunnTest(Percent_Survival ~ Parent_pCO2,
                data = df_F3,
              method="bh")
```